# Sample selection

We select sampling units from the table `mhq_watersurfaces_populationunits`.
The selection is based on the value `grts_ranking_draw`.

We also check how many of the selected sampling units have already been measured in the first monitoring cycle and how many of the measured populations are not selected.



```{r}

root_samplingframe <- "../../habitatwatersurfaces/output" 

watersurfaces_populationunits <- read_vc(root = root_samplingframe, file = "mhq_watersurfaces_populationunits")

assessments <- read_vc(root = root_samplingframe, file = "mhq_watersurfaces_assessments")

measurements <- read_vc(root = root_samplingframe, file = "mhq_watersurfaces_measurements")

measurements_summary <- measurements %>%
  filter(!is.na(type)) %>%
  group_by(polygon_id, type) %>%
  summarise(measurement_type = str_c(measurement_type, collapse = " + ")) %>%
  ungroup()


check <- read_watersurfaces_hab()
check <- check$watersurfaces_types
```


```{r}
strata <- watersurfaces_types_2021 %>%
  distinct(polygon_id, area_class, sac) 

strata_watersurfaces_extra <- read_watersurfaces() %>%
  filter(polygon_id %in% assessments$polygon_id) %>%
  filter(! polygon_id %in% strata$polygon_id) %>%
  mutate(area_ha = drop_units(st_area(geom))/10000,
         area_polygon_ha = area_ha,
         area_class = ifelse(area_ha <= 1, "Opp <= 1 ha",
                             ifelse(area_ha <= 5, "1 ha < Opp <= 5 ha",
                                    ifelse(area_ha < 50, "5 ha < Opp < 50 ha", ">= 50 ha"))),
         area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp < 50 ha", ">= 50 ha"))) %>%
  st_join(sac_sf) %>%
  mutate(sac = !is.na(sac_name)) %>%
  st_drop_geometry() %>%
  distinct(polygon_id, area_class, sac)

strata <- bind_rows(strata,
                    strata_watersurfaces_extra)

populationunits_details <- watersurfaces_populationunits %>%
  left_join(strata, by = "polygon_id") %>%
  left_join(measurements_summary, by = c("polygon_id", "type")) %>%
  mutate(measured_cycle1 = !is.na(measurement_type)) %>%
  mutate(main_type = str_sub(type, 1, 4),
         type = ifelse(type == "3130", "3130_aom", as.character(type)))

# rank_main_type <- populationunits_details %>%
#   distinct(polygon_id, main_type, sac, area_class, grts_ranking_draw) %>%
#   group_by(main_type, sac, area_class) %>%
#   mutate(rank_main_type = rank(grts_ranking_draw)) %>%
#   ungroup() %>%
#   select(polygon_id, main_type, rank_main_type)
# 
# populationunits_details <- populationunits_details %>%
#   left_join(rank_main_type, by = c("polygon_id", "main_type"))

```

```{r}
samplesize_main_type <- samplesize_alt %>%
  filter(version == "watersurfaces_hab_2021") %>%
  distinct( main_type, sac, area_class, samplesize_main_type_alt1) %>%
  mutate(samplesize_main_type_alt1 = round(samplesize_main_type_alt1))

samplesize_subtype <- samplesize_alt %>%
  filter(version == "watersurfaces_hab_2021") %>%
  distinct(type, sac, area_class, samplesize_subtype_alt1) %>%
  mutate(samplesize_subtype_alt1 = round(samplesize_subtype_alt1)) %>%
  filter(!is.na(samplesize_subtype_alt1))

samplesize_type_total <- samplesize_alt %>%
  filter(version == "watersurfaces_hab_2021") %>%
  distinct(type, sac, area_class, samplesize_total_alt1) %>%
  mutate(samplesize_total_alt1 = round(samplesize_total_alt1))

selection_type <- populationunits_details %>%
  left_join(samplesize_type_total, by = c("type", "sac", "area_class")) %>%
  # left_join(samplesize_main_type, by = c("main_type", "sac", "area_class")) %>%
  # left_join(samplesize_subtype, by = c("type", "sac", "area_class")) %>%
   group_by(type, sac, area_class) %>%
  mutate(rank_subtype = rank(grts_ranking_draw)) %>%
  ungroup() %>%
  mutate(selected = rank_subtype <= samplesize_total_alt1,
         reserve = (!selected) & (rank_subtype <= samplesize_total_alt1 + 10)) %>%
  group_by(polygon_id) %>%
  mutate(type_all = str_c(type, collapse = " + ")) %>%
  ungroup()
  # group_by(type, sac, area_class) %>%
  # mutate(n_selected_type = sum(selected),
  #        rank_subtype = rank(grts_ranking_draw)) %>%
  # ungroup() %>%
  # mutate(n_extra_subtype = pmax(0, samplesize_subtype_alt1 -n_selected_type)) %>%
  # mutate(selected_extra = (rank_subtype > n_selected_type) & (rank_subtype <= (n_selected_type + n_extra_subtype)),
  #        selected = ifelse(selected, TRUE,
  #                          ifelse(selected_extra & !is.na(selected_extra), TRUE, 
  #                                 FALSE)))

selection_type_overview <- selection_type %>%
  group_by(main_type, type, sac, area_class, selected, measured_cycle1) %>%
  summarise(n = n_distinct(polygon_id)) %>%
  ungroup() %>%
  mutate(names = ifelse(selected,
                      ifelse(measured_cycle1, "selected_measured", "selected_notmeasured"),
                      ifelse(measured_cycle1, "notselected_measured", "notselected_notmeasured"))) %>%
  select(-selected, -measured_cycle1) %>%
  pivot_wider(names_from = names, values_from = n, values_fill = 0) %>%
  mutate(selected_total = selected_measured + selected_notmeasured,
         total = selected_total + notselected_measured + notselected_notmeasured,
         total_measured = selected_measured + notselected_measured) %>%
  left_join(samplesize_type_total, by = c("type", "sac", "area_class")) 
```

Table \@ref(tab:resultalt1) shows the results per type, area_class and sac for sample size alternative 1 with n_sac = 80:
+ the number of selected sampling units 
+ the number of selected sampling units that haven been measured (the first number within parentheses)
+ the number of population units (selected and not selected) that haven been measured (the second number within parentheses)



```{r resultalt1}
selection_type_overview %>%
  mutate(show_n = str_c( selected_total, " (", selected_measured, "/", total_measured, ")")) %>%
  select(type, sac, area_class, show_n) %>%
  pivot_wider(names_from = area_class, values_from = show_n, values_fill = "0") %>%
  kable(caption = "Sample size for altenative 1 with n_sac = 80 (number of selected sampling units that have been measured in the first monitoring cycle/ total number of measured watersurfaces in the first monitoring cycle)") %>%
  kable_styling()

```


```{r}
watersurfaces_sample <- watersurfaces_area_sf %>%
  select(polygon_id, sac, area_class) %>%
  st_centroid() %>%
  left_join(select(selection_type, main_type, type, type_all, polygon_id, selected, measured_cycle1, grts_ranking, grts_ranking_draw), by = "polygon_id") %>%
  filter(selected)

flanders <- read_admin_areas(dsn = "flanders")

plot1 <- ggplot() +
  geom_sf(data = flanders, fill = "white") +
  geom_sf(data = watersurfaces_sample, aes(colour = area_class), alpha = 0.6) +
  labs(colour = "Oppervlakteklasse") +
  theme(legend.text = element_text(size = 30),
        legend.title = element_text(size = 30))


ggsave(filename = "../output/steekproef_watervlakken.png", plot = plot1, width = 8, height = 3)

```




```{r}

watersurfaces_show <- watersurfaces_area_sf %>%
  select(polygon_id, sac, area_class) %>%
  st_centroid() %>%
  left_join(select(selection_type, main_type, type, type_all, polygon_id, selected, measured_cycle1, grts_ranking, grts_ranking_draw)) %>%
  st_transform(crs = 4326) %>%
  mutate(color_fill_show = ifelse(selected, "blue", "red"),
         color_stroke_show = ifelse(measured_cycle1, "yellow", "grey"),
         label_show = str_c("id: ",polygon_id, 
                            "; type: ", type_all,
                            "; sac: ", sac,
                            "; area_class: ", area_class),
         type = as.character(type))
  
```

The map below gives an overview of the sample selection:

+ selected sampling units (blue circles)
+ population units that have not been selected (red circles)
+ population units that have been measured in the first monitoring cycle (circle with yellow line)
+ population units that have not been measured in the first monitoring cycle (circle with grey line)

By clicking on a circle marker in the map you can see the polygon_id, type, area_class and sac.

```{r}

sd <- SharedData$new(watersurfaces_show)

bscols(filter_select("type", "Select habitat type", sharedData = sd, ~type, allLevels = FALSE),
       filter_select("area_class", "Area class", sharedData = sd, ~area_class, allLevels = FALSE),
        filter_checkbox("sac", "Within sac", sharedData = sd, ~sac)
       )

bscols(filter_checkbox("measured_cycle1", "Measured in cycle 1", sharedData = sd, ~measured_cycle1),
       filter_checkbox("selected", "Selected sampling unit", sharedData = sd, ~selected)
       )

bscols(widths = c(8, 4),
        leaflet(sd) %>%
        addTiles() %>%
        addCircleMarkers(fillColor = ~color_fill_show, color =  ~color_stroke_show, fillOpacity = 0.4, opacity = 0.7,
                         popup = ~label_show),
       plot_ly(sd, x = ~measured_cycle1) %>%
        add_histogram(color = ~selected,
                      colors = c("red", "blue")) %>%
        layout(barmode = "stack",
               legend=list(title=list(text='<b> Selected sampling unit </b>'))) %>%
      layout(legend = list(x = 200, y = 0.5))
)




```

# Include more sampling units that were measured in the first monitoring cycle?

It could be an option to include in the sample more watersurfaces that have been measured in the first monitoring cycle.

Advantages are:

+ more paired observations and therefore a higher precision for trend estimates
+ the sample contains more watersurfaces that have already been visited and therefore the presence of the target type is known (with high probability)

The disadvantage is that:

+ the sample is less representative for the current distribution of the habitat types and there fore estimates of status and trend could be biased (and since there is a large difference between the original sampling frame and the current one, this bias could be significant)

```{r}
check_measurements <- populationunits_details %>%
  filter(str_detect(source, pattern = "original sample frame|assessments monitoring cycle 1")) %>%
  group_by(type, sac, area_class) %>%
  mutate(rank_type = rank(grts_ranking_draw),
         n_measured = sum(measured_cycle1)) %>%
  ungroup() %>%
  mutate(selection_correct = rank_type <= n_measured) %>%
  group_by(type, sac, area_class, n_measured) %>%
  summarise(n_selection_correct = sum(selection_correct * measured_cycle1)) %>%
  ungroup()
  
```

If we would include more measured watersurfaces in the current sample, we should only take those watersurfaces that would have been selected based on the original sampling frame.

In Table \@ref(tab:tabmeasurements) we show: 

+ the measured watersurfaces hat would have been selected based on the original sampling frame + the total number of measured watersurfaces

This evaluation is based on the current design: we take a sample for all types (except 3110) and all area classes (except > 50 ha).

```{r tabmeasurements}
check_measurements %>%
  mutate(show_n = str_c(n_selection_correct,"/", n_measured)) %>%
  select(type, sac, area_class, show_n) %>%
  pivot_wider(names_from = area_class, values_from = show_n, values_fill = "0") %>%
  kable(caption = "Number of measured sampling units that belong to the sample based on the original sampling frame/ Total number of measured sampling units") %>%
  kable_styling()

```

















