# Sample frame

## Sample frame used in Westra et al. (2014)


```{r}
versie <- "20140324"

watersurfaces_hab_polygons_2014_orig <- read_sf("../../habitatwatersurfaces/data/watersurfaces_2014/BWK_plassen_versie 20140324.shp")

watersurfaces_hab_polygons_2014 <- watersurfaces_hab_polygons_2014_orig %>%
  mutate(sac = (SBZH == 1) & (!is.na(SBZH))) %>%
  select(pol_id = Pol_ID, sac)

check <- nrow(watersurfaces_hab_polygons_2014) == n_distinct(watersurfaces_hab_polygons_2014$pol_id)

watersurfaces_hab_types_2014_orig <- read.table("../../habitatwatersurfaces/data/watersurfaces_2014/BWK_hab_aq.txt", header=TRUE)

watersurfaces_hab_types_2014 <- watersurfaces_hab_types_2014_orig %>%
  select(pol_id = Pol_ID, shape_area = Shape_Area, code, phab) %>%
  filter(str_detect(code, "gh", negate = TRUE)) %>%
  mutate(code = ifelse(code == "3130", "3130_aom", code),
         patch_area = shape_area * phab/100,
         area_ha = patch_area/10000,
         area_polygon_ha = shape_area/10000,
         area_class = ifelse(area_ha <= 1, "Opp <= 1 ha",
                             ifelse(area_ha <= 5, "1 ha < Opp <= 5 ha",
                                    ifelse(area_ha < 50, "5 ha < Opp < 50 ha", ">= 50 ha"))),
         area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp < 50 ha", ">= 50 ha"))) %>%
  left_join(st_drop_geometry(watersurfaces_hab_polygons_2014), by = "pol_id") %>%
  rename(type = code) %>%
  mutate(version = "habitatmap_2014")

```


### Habitatmap 2020

```{r}

sac_sf <- read_admin_areas(dsn = "sac") %>%
  group_by(sac_name) %>%
  summarise(n_pol = n()) %>%
  ungroup() %>%
  select(sac_name)

habitatmap_2020 <- read_habitatmap_stdized()

watersurfaces_hab_polygons_2020 <- habitatmap_2020$habitatmap_polygons %>%
  filter(str_detect(description_orig, "31")) %>%
  mutate(area_polygon_ha = drop_units(st_area(geom)) / 10000) %>%
  select(polygon_id, area_polygon_ha) %>%
  st_join(sac_sf, largest = TRUE) %>%
  mutate(sac = !is.na(sac_name)) %>%
  st_drop_geometry()

watersurfaces_hab_types_2020 <- habitatmap_2020$habitatmap_types %>%
  filter(str_sub(type,1, 2) == "31") %>%
  left_join(watersurfaces_hab_polygons_2020, by = "polygon_id") %>%
  mutate(area_ha = area_polygon_ha * phab/100,
         area_class = ifelse(area_ha <= 1, "Opp <= 1 ha",
                             ifelse(area_ha <= 5, "1 ha < Opp <= 5 ha",
                                    ifelse(area_ha < 50, "5 ha < Opp < 50 ha", ">= 50 ha"))),
         area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp < 50 ha", ">= 50 ha")),
         type = ifelse(type == "3130", "3130_aom", as.character(type))) %>%
  mutate(version = "habitatmap_2020")
```


### Watersurfaces_hab

```{r}
watersurfaces_hab <- read_watersurfaces_hab()

filepath <-  file.path(fileman_up("n2khab_data"), "20_processed/watersurfaces_hab/watersurfaces_hab.gpkg")

check_version <- n2khab::xxh64sum(filepath) == "5792b496a94d0524"



watersurfaces_area <- watersurfaces_hab$watersurfaces_polygons %>%
  mutate(area_ha = drop_units(st_area(geom))/10000,
         area_polygon_ha = area_ha,
         area_class = ifelse(area_ha <= 1, "Opp <= 1 ha",
                             ifelse(area_ha <= 5, "1 ha < Opp <= 5 ha",
                                    ifelse(area_ha < 50, "5 ha < Opp < 50 ha", ">= 50 ha"))),
         area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp < 50 ha", ">= 50 ha"))) %>%
  st_join(sac_sf) %>%
  mutate(sac = !is.na(sac_name)) %>%
  st_drop_geometry()

watersurfaces_types_2021 <- watersurfaces_hab$watersurfaces_types %>%
  filter(type != "rbbah") %>%
  left_join(watersurfaces_area, by = "polygon_id") %>%
  mutate(type = ifelse(type == "3130", "3130_aom", as.character(type))) %>%
  mutate(version = "watersurfaces_hab_2021")

watersurfaces_types <- watersurfaces_types_2021 %>%
  bind_rows(watersurfaces_hab_types_2014) %>%
  bind_rows(watersurfaces_hab_types_2020)

```

## Compare sample frame versions

```{r}
watersurfaces_hab_distribution <- watersurfaces_types %>%
  group_by(version, type, sac, area_class) %>%
  summarise(n_watersurfaces = n(),
            area_ha = sum(area_ha)) %>%
  ungroup() %>%
  group_by(version, type, sac) %>%
  mutate(total_n_sac = sum(n_watersurfaces),
         total_area_sac = sum(area_ha)) %>%
  ungroup() 

```

```{r}
watersurfaces_hab_distribution %>%
  select(type, version, everything()) %>%
  mutate(n_area_show = str_c(n_watersurfaces, " (", round(area_ha, 1), " ha)"),
         total = str_c(total_n_sac, " (", round(total_area_sac, 1), " ha)"),
         .keep = "unused") %>%
  spread(key = area_class, value = n_area_show, fill = 0) %>%
  kable() %>%
  kable_styling() %>%
  collapse_rows(c(1,2))
```




```{r, fig.height= 9}
watersurfaces_types %>%
  ggplot(aes(x = area_class, fill = version)) +
  geom_bar(position = "dodge") +
  facet_wrap(~type, ncol = 2, scales = "free_x") +
  coord_flip() +
  theme(legend.position = "bottom")
```

```{r, fig.height= 9}
watersurfaces_types %>%
  group_by(version, type, area_class) %>%
  summarise(area_ha = sum(area_ha)) %>%
  ungroup() %>%
  ggplot(aes(x = area_class, y = area_ha , fill = version, shape = version)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~type, ncol = 2, scales = "free_x") +
  coord_flip() +
  theme(legend.position = "bottom")
```
```{r, fig.height= 9}
watersurfaces_types %>%
  group_by(version, type, area_class) %>%
  summarise(area_polygon_ha = sum(area_polygon_ha)) %>%
  ungroup() %>%
  ggplot(aes(x = area_class, y = area_polygon_ha , fill = version, shape = version)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~type, ncol = 2, scales = "free_x") +
  coord_flip() +
  theme(legend.position = "bottom")
```


```{r}


area_type_reporting2019 <- read_csv2("../../habitatwatersurfaces/data/type_area_reporting2019.csv") %>%
  filter(str_sub(type, 1, 2) == "31") %>%
  mutate(version = "rapportage_2019",
         area = round(area,1)) %>%
  select(type, area_ha = area, version) 

watersurfaces_types %>%
  group_by(type, version) %>%
  summarise(area_ha = round(sum(area_ha), 1),
            n_watersurfaces = n(),
            area_polygon_ha = round(sum(area_polygon_ha), 1)) %>%
  ungroup() %>%
  bind_rows(area_type_reporting2019) %>%
  arrange(type, version) %>%
  kable() %>%
  kable_styling() %>%
  collapse_rows()
```



# Sample frame

## Original version

+ sample for 3130_aom, 3150, 3160 in area classes `Opp <= 1 ha` and  `1 ha < Opp <= 5 ha`
  + sample size = 40 for Flanders
  + extra sampling units within sac until 80 sampling units are selected
  + finite population correction
  + sample size equally divided over both area classes
+ selection of all watersurfaces for other area classes and other habitat types

### Westra et al. (2014)

```{r}
samplesize_original <- read_table2("../../habitatwatersurfaces/data/watersurfaces_2014/Steekproefgrootte_plassen_versie20140326.txt") %>%
  rename(sac = "SBZ-H") %>%
  mutate(sac = sac == 1) %>%
  select(type = habt, sac, area_class_min = min_opp, area_class_max = max_opp, samplesize = n) 

```


```{r}
samplesize_original %>%
  filter(type %in% c("3130_aom", "3150", "3160")) %>%
  mutate(area_class = ifelse(area_class_max == 1, "Opp <= 1 ha",
                             ifelse(area_class_max == 5, "1 ha < Opp <= 5 ha",
                                    "5 ha < Opp <= 50 ha")),
         area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp <= 50 ha"))) %>%
  select(-area_class_max, -area_class_min) %>%
  spread(key = area_class, value = samplesize, fill = 0) %>%
  kable() %>%
  kable_styling()
```

Hmmm... still different than in the Westra et al. report...


### Reproduce samplesize from sample frame

```{r}
n_fland <- 40
n_sac <- 80

samplesize <- watersurfaces_hab_distribution %>%
  group_by(version, type, area_class) %>%
  mutate(total_area_class = sum(n_watersurfaces)) %>%
  ungroup() %>%
  group_by(version, type) %>%
  mutate(total = sum(n_watersurfaces)) %>%
  ungroup() %>%
  mutate(take_sample = total >= 50 & area_class %in% c("Opp <= 1 ha", "1 ha < Opp <= 5 ha")) %>%
  group_by(version, type, take_sample) %>%
  mutate(n_watersurfaces_sample = sum(n_watersurfaces * take_sample)) %>%
  ungroup() %>% 
  group_by(version, type, take_sample, sac) %>%
  mutate(n_watersurfaces_sample_sac = sum(n_watersurfaces * take_sample)) %>%
  ungroup() %>%
  mutate(n_finite_fland = n_watersurfaces_sample * n_fland / (n_fland + (n_watersurfaces_sample - 1)),
         n_finite_sac = n_watersurfaces_sample_sac * n_sac / (n_sac + (n_watersurfaces_sample_sac - 1)),
         prop_sac = n_watersurfaces_sample_sac/n_watersurfaces_sample, 
         samplesize = ifelse(!take_sample, n_watersurfaces,
                             ifelse(sac, n_finite_sac/2, n_finite_fland * prop_sac/2)),
         samplesize = pmin(samplesize, n_watersurfaces))

```

```{r}
samplesize_overview <- samplesize %>%
  filter(version %in% c("habitatmap_2014", "watersurfaces_hab_2021")) %>%
  select(type, version, sac, area_class, samplesize, n_watersurfaces) %>%
  mutate(samplesize_show = str_c(round(samplesize), "/", n_watersurfaces),
         .keep = "unused") %>%
  spread(key = area_class, value = samplesize_show, fill = 0) %>%
  arrange(type, version) %>%
  mutate(scenario = "original")

samplesize_overview %>%
  kable() %>%
  kable_styling() %>%
  collapse_rows(c(1,2))
  
```


```{r}
samplesize_total <- samplesize %>%
  filter(version %in% c("habitatmap_2014", "watersurfaces_hab_2021")) %>%
  group_by(version, area_class) %>%
  summarise(samplesize = sum(samplesize),
            n_watersurfaces = sum(n_watersurfaces)) %>%
  ungroup() %>%
  mutate(samplesize_show = str_c(round(samplesize), "/", n_watersurfaces),
         .keep = "unused") %>%
  spread(key = area_class, value = samplesize_show, fill = 0) %>%
  arrange(version) %>%
  mutate(scenario = "original")
```

```{r}
samplesize_total %>%
  kable() %>%
  kable_styling()
```



### Alternative scenario's

```{r}
n_fl <- 80
n_sac <- 80
n_subtype <- 40


samplesize_fl <- samplesize %>%
  mutate(main_type = str_sub(type, 1, 4)) %>%
  group_by(version, main_type, area_class, sac) %>%
  summarise(area_ha_sac = sum(area_ha),
            n_watersurfaces_sac = sum(n_watersurfaces)) %>%
  ungroup() %>%
  group_by(version, main_type, area_class) %>%
  mutate(n_watersurfaces = sum(n_watersurfaces_sac),
         area_ha = sum(area_ha_sac)) %>%
  ungroup() %>%
  mutate(weight1 = log(area_ha + 1),
          weight2 = area_ha ^ (1/2)) %>%
  group_by(version, main_type, sac) %>%
  mutate(total_area_ha = sum(area_ha * (area_class != ">= 50 ha")),
         total_watersurfaces = sum(n_watersurfaces * (area_class != ">= 50 ha")),
         total_weight1 = sum(weight1 * (area_class != ">= 50 ha")),
         total_weight2 = sum(weight2 * (area_class != ">= 50 ha"))) %>%
  ungroup() %>%
  mutate(n_finite_fland = total_watersurfaces * n_fl / (n_fl + (total_watersurfaces - 1))) %>%
  mutate(samplesize_fl_alt1 = ifelse(area_class != ">= 50 ha", n_finite_fland * weight1 / total_weight1, n_watersurfaces),
         samplesize_fl_alt2 = ifelse(area_class != ">= 50 ha", n_finite_fland * weight2 / total_weight2, n_watersurfaces),
         samplesize_fl_alt3 = ifelse(area_class != ">= 50 ha", n_finite_fland /3, n_watersurfaces)) %>%
  mutate(samplesize_fl_alt1 = pmin(samplesize_fl_alt1 * n_watersurfaces_sac/n_watersurfaces, n_watersurfaces),
         samplesize_fl_alt2 = pmin(samplesize_fl_alt2 * n_watersurfaces_sac/n_watersurfaces, n_watersurfaces),
         samplesize_fl_alt3 = pmin(samplesize_fl_alt3 * n_watersurfaces_sac/n_watersurfaces, n_watersurfaces)) 


samplesize_sac <- samplesize %>%
  filter(sac) %>%
  mutate(main_type = str_sub(type, 1, 4)) %>%
  group_by(version, sac, main_type, area_class) %>%
  summarise(area_ha = sum(area_ha),
            n_watersurfaces = sum(n_watersurfaces)) %>%
  ungroup() %>%
  mutate(weight1 = log(area_ha + 1),
          weight2 = area_ha ^ (1/2)) %>%
  group_by(version, main_type) %>%
  mutate(total_area_ha = sum(area_ha * (area_class != ">= 50 ha")),
         total_watersurfaces = sum(n_watersurfaces),
         total_weight1 = sum(weight1 * (area_class != ">= 50 ha")),
         total_weight2 = sum(weight2 * (area_class != ">= 50 ha"))) %>%
  ungroup() %>%
  mutate(n_finite_sac = total_watersurfaces * n_sac / (n_sac + (total_watersurfaces - 1))) %>%
  mutate(samplesize_sac_alt1 = ifelse(area_class != ">= 50 ha", n_finite_sac * weight1 / total_weight1, n_watersurfaces),
         samplesize_sac_alt2 = ifelse(area_class != ">= 50 ha", n_finite_sac * weight2 / total_weight2, n_watersurfaces),
         samplesize_sac_alt3 = ifelse(area_class != ">= 50 ha", n_finite_sac /3, n_watersurfaces)) %>%
  mutate(samplesize_sac_alt1 = pmin(samplesize_sac_alt1, n_watersurfaces),
         samplesize_sac_alt2 = pmin(samplesize_sac_alt2, n_watersurfaces),
         samplesize_sac_alt3 = pmin(samplesize_sac_alt3, n_watersurfaces))

samplesize_subtype <- samplesize %>%
  mutate(main_type = str_sub(type, 1, 4)) %>%
  filter(main_type == "3130") %>%
  group_by(version, type, area_class, sac) %>%
  summarise(n_watersurfaces_sac = sum(n_watersurfaces),
            area_ha_sac = sum(area_ha)) %>%
  ungroup() %>%
  group_by(version, type, area_class) %>%
  mutate(n_watersurfaces = sum(n_watersurfaces_sac),
         area_ha = sum(area_ha_sac)) %>%
  ungroup() %>%
  mutate(weight1 = log(area_ha + 1),
          weight2 = area_ha ^ (1/2)) %>%
  group_by(version, type, sac) %>%
  mutate(total_area_ha = sum(area_ha * (area_class != ">= 50 ha")),
         total_watersurfaces = sum(n_watersurfaces),
         total_weight1 = sum(weight1 * (area_class != ">= 50 ha")),
         total_weight2 = sum(weight2 * (area_class != ">= 50 ha"))) %>%
  ungroup() %>%
  mutate(n_finite_subtype = total_watersurfaces * n_subtype / (n_subtype + (total_watersurfaces - 1))) %>%
  mutate(samplesize_subtype_alt1 = ifelse(area_class != ">= 50 ha", n_finite_subtype * weight1 / total_weight1, n_watersurfaces),
         samplesize_subtype_alt2 = ifelse(area_class != ">= 50 ha", n_finite_subtype * weight2 / total_weight2, n_watersurfaces),
         samplesize_subtype_alt3 = ifelse(area_class != ">= 50 ha", n_finite_subtype /3, n_watersurfaces)) %>%
  mutate(samplesize_subtype_alt1 = pmin(samplesize_subtype_alt1 * n_watersurfaces_sac/n_watersurfaces, n_watersurfaces_sac),
         samplesize_subtype_alt2 = pmin(samplesize_subtype_alt2 * n_watersurfaces_sac/n_watersurfaces, n_watersurfaces_sac),
         samplesize_subtype_alt3 = pmin(samplesize_subtype_alt3 * n_watersurfaces_sac/n_watersurfaces, n_watersurfaces_sac))

samplesize_alt <- samplesize %>%
  mutate(main_type = str_sub(type, 1, 4)) %>%
  select(version, main_type, type, sac, area_class, area_ha, n_watersurfaces) %>%
  group_by(version, main_type) %>%
  mutate(n_watersurfaces_sample = sum(n_watersurfaces)) %>%
  ungroup() %>%
  left_join(select(samplesize_fl, sac, version, main_type, area_class, samplesize_fl_alt1, samplesize_fl_alt2, samplesize_fl_alt3), by = c("version", "main_type", "area_class", "sac")) %>%
  left_join(select(samplesize_sac, sac, version, main_type, area_class, samplesize_sac_alt1, samplesize_sac_alt2, samplesize_sac_alt3), by = c("version", "main_type", "area_class", "sac")) %>%
  group_by(version, main_type, sac, area_class) %>%
  mutate(n_watersurfaces_main_type = sum(n_watersurfaces)) %>%
  ungroup() %>%
  mutate(samplesize_main_type_alt1 = ifelse(area_class == ">= 50 ha" | n_watersurfaces_sample < 30, n_watersurfaces,
                                            ifelse(sac, samplesize_sac_alt1, samplesize_fl_alt1)),
         samplesize_main_type_alt2 = ifelse(area_class == ">= 50 ha" | n_watersurfaces_sample < 30, n_watersurfaces,
                                            ifelse(sac, samplesize_sac_alt2, samplesize_fl_alt2)),
         samplesize_main_type_alt3 = ifelse(area_class == ">= 50 ha" | n_watersurfaces_sample < 30, n_watersurfaces,
                                            ifelse(sac, samplesize_sac_alt3, samplesize_fl_alt3)),
         samplesize_type_alt1 = samplesize_main_type_alt1 * n_watersurfaces / n_watersurfaces_main_type,
         samplesize_type_alt2 = samplesize_main_type_alt2 * n_watersurfaces / n_watersurfaces_main_type,
         samplesize_type_alt3 = samplesize_main_type_alt3 * n_watersurfaces / n_watersurfaces_main_type,
         ) %>%
  left_join(select(samplesize_subtype, sac, version, type, area_class, samplesize_subtype_alt1, samplesize_subtype_alt2, samplesize_subtype_alt3), by = c("version", "type", "area_class", "sac")) %>%
  mutate(samplesize_total_alt1 = round(pmax(samplesize_type_alt1, samplesize_subtype_alt1, na.rm = TRUE)),
         samplesize_total_alt2 = round(pmax(samplesize_type_alt2, samplesize_subtype_alt2, na.rm = TRUE)),
         samplesize_total_alt3 = round(pmax(samplesize_type_alt3, samplesize_subtype_alt3, na.rm = TRUE)))

```

```{r}
samplesize_alt_total <- samplesize_alt %>%
  select(version, type, area_class, sac, n_watersurfaces, starts_with("samplesize_total")) %>%
  pivot_longer(starts_with("samplesize_total"), names_to = "scenario", names_prefix = "samplesize_total_", values_to = "samplesize") %>% group_by(scenario, version, area_class) %>%
  summarise(samplesize = sum(samplesize),
            n_watersurfaces = sum(n_watersurfaces)) %>%
  ungroup() %>%
  mutate(samplesize_show = str_c(round(samplesize), "/", n_watersurfaces),
         .keep = "unused") %>%
  spread(key = area_class, value = samplesize_show, fill = 0) %>%
  bind_rows(samplesize_total) %>%
  arrange(version, scenario) 
```
```{r}
samplesize_alt_total %>%
  kable() %>%
  kable_styling()
```
```{r}
samplesize_alt_overview <- samplesize_alt%>%
  select(version, type, area_class, sac, n_watersurfaces, starts_with("samplesize_total")) %>%
  pivot_longer(starts_with("samplesize_total"), names_to = "scenario", names_prefix = "samplesize_total_", values_to = "samplesize") %>%
  filter(version %in% c("watersurfaces_hab_2021")) %>%
  mutate(samplesize_show = str_c(round(samplesize), "/", n_watersurfaces),
         .keep = "unused") %>%
  spread(key = area_class, value = samplesize_show, fill = 0) %>%
  select(version, scenario, everything()) %>%
  bind_rows(filter(samplesize_overview, version == "habitatmap_2014")) %>%
  arrange(type, scenario) 

samplesize_alt_overview %>%
  kable() %>%
  kable_styling() %>%
  collapse_rows(c(1,2), target = 2)
  
````

