
# Read original mhq sample for watersurfaces

## Selected sampling units in 2014

In 2014 the an initial sample was drawn based in the habitatmap version 2014.
The sampling unit corresponds to a polygon of the habitatmap.
A GRTS ranking was asigned to every polygon based on [GRTSmaster_habitats](https://zenodo.org/record/2682323).
The centroid of the polygon was used to link the polygon to a GRTS raster cell.
So we can regard the centroid of the polygon as the 'reference point' of the sampling unit

The files can be found in [this google drive folder](https://drive.google.com/drive/folders/1j-CbsuF6vTQJ3weT6gJWTV1nVhxWmfX7).


```{r}
# shapefile with polygons
sample_2014_orig <- read_sf("../data/watersurfaces_2014/steekproef2014.shp", crs = 31370)

sample_2014_sf <- sample_2014_orig %>%
  select(pol_id = Pol_ID, sac = SBZH, type_target = doelhabt) %>%
  mutate(sac = ifelse(is.na(sac), 0, sac)) %>%
  group_by(pol_id, sac) %>%
  arrange(type_target) %>%
  summarise(type_target_all = str_c(type_target, collapse = ";")) %>%
  ungroup() %>%
  mutate(pol_id = as.character(pol_id))

# text file with sampling units
sample_3130_aom <- read.table("../data/watersurfaces_2014/meetnet3130_aom_20140324.txt", ) 
sample_3130_na <- read.table("../data/watersurfaces_2014/meetnet3130_na_20140324.txt") 
sample_3140 <- read.table("../data/watersurfaces_2014/meetnet3140_20140324.txt") %>%
  mutate(code = as.character(code))
sample_3150 <- read.table("../data/watersurfaces_2014/meetnet3150_20140324.txt")  %>%
  mutate(code = as.character(code))
sample_3160 <- read.table("../data/watersurfaces_2014/meetnet3160_20140324.txt")%>%
  mutate(code = as.character(code))

sample_2014 <- bind_rows(sample_3130_aom,
                         sample_3130_na,
                         sample_3140,
                         sample_3150,
                         sample_3160) %>%
  filter(steekproef == 1) %>%
  select(pol_id = Pol_ID, sac = SBZH, type_target = code, phab, area_class = Opp_klasse, grts_ranking = Ranking, ranking_rel = rangnr) %>%
  mutate(set = "selection2014",
    area_class = ifelse(area_class == "(0,1e+04]", "Opp <= 1 ha",
                             ifelse(area_class == "(1e+04,5e+04]", "1 ha < Opp <= 5 ha",
                                    ifelse(area_class == "(5e+04,1e+05]", "5 ha < Opp < 50 ha",
                                           ifelse(area_class == "(1e+05,2.5e+05]", "5 ha < Opp < 50 ha",
                                                  ifelse(area_class == "(2.5e+05,5e+05]", "5 ha < Opp < 50 ha", NA))))),
    area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp < 50 ha"))) %>%
  mutate(pol_id = as.character(pol_id))

sample_2014_ranking <- sample_2014 %>%
  distinct(pol_id, grts_ranking)

sample_2014_sf <- sample_2014_sf %>%
  left_join(sample_2014_ranking, by = "pol_id") %>%
  mutate(set = "selection2014")

check_pol_id <- all.equal((sample_2014_sf %>%
                      arrange(pol_id))$pol_id,
                    (sample_2014 %>%
                      arrange(pol_id))$pol_id %>%
                      unique())
```

The centroids of the original sampling units were also saved as shape files. 

```{r}
# read original refpoint data
grts_master <- read_GRTSmh()

refpoint_3130_aom <- read_sf("../data/watersurfaces_2014/meetnet_3130_aom_versie20140324.shp") %>%
  st_drop_geometry() %>%
  select(pol_id = Pol_ID, x_centroid = centroidX, y_centroid = centroidY) 

refpoint_3130_na <- read_sf("../data/watersurfaces_2014/meetnet_3130_na_versie20140324.shp") %>%
  st_drop_geometry() %>%
  select(pol_id = Pol_ID, x_centroid = centroidX, y_centroid = centroidY) 

refpoint_3140 <- read_sf("../data/watersurfaces_2014/meetnet_3140_versie20140324.shp") %>%
  st_drop_geometry() %>%
  select(pol_id = Pol_ID, x_centroid = centroidX, y_centroid = centroidY) 

refpoint_3150 <- read_sf("../data/watersurfaces_2014/meetnet_3150_versie20140324.shp") %>%
  st_drop_geometry() %>%
  select(pol_id = Pol_ID, x_centroid = centroidX, y_centroid = centroidY)

refpoint_3160 <- read_sf("../data/watersurfaces_2014/meetnet_3160_versie20140324.shp") %>%
  st_drop_geometry() %>%
  select(pol_id = Pol_ID, x_centroid = centroidX, y_centroid = centroidY)

sample_2014_refpoint <- bind_rows(refpoint_3130_aom, refpoint_3130_na, refpoint_3140, refpoint_3150, refpoint_3160) %>%
  unique() %>%
  mutate(x_refpoint = x_centroid,
         y_refpoint = y_centroid) %>%
  st_as_sf(coords = c("x_centroid", "y_centroid"), crs = 31370) %>%
  mutate(grts_ranking = grts_master[as(., "Spatial")],
         pol_id = as.character(pol_id))

```


### Reproduce grts_ranking based on original reference points

For 2 sampling units the originally determined grts-ranking differs from the reproduced grts-ranking:

+ the grts-ranking were switched for the polygons with pol_id 5083 and 16115 due to a manual error.

We will correct the grts-ranking and use the variable grts_ranking_draw to store the original grts-ranking that was used to draw the sample.

```{r}
check_ranking <- sample_2014_ranking %>%
  left_join(st_drop_geometry(sample_2014_refpoint), suffix = c("_orig", "_check"), by = "pol_id") %>%
  filter(grts_ranking_orig != grts_ranking_check)
  
check_ranking %>%
  select(-x_refpoint, -y_refpoint) %>%
  kable() %>%
  kable_styling()
```

### Reproduce centroids

We also check if we can reproduce the centroids, as a different location of the centroid can result in a differncet grs-ranking.

```{r}

sample_2014_ranking_centroid <- sample_2014_sf %>%
  st_centroid() %>%
  mutate(grts_ranking_centroid = grts_master[as(., "Spatial")])

sample_2014_ranking_centroid <- sample_2014_ranking_centroid %>%
  select(-grts_ranking) %>%
  mutate(x_centroid = st_coordinates(geometry)[, 1],
         y_centroid = st_coordinates(geometry)[, 2])

check_centroid <- sample_2014_ranking_centroid %>%
  st_drop_geometry() %>%
  left_join(st_drop_geometry(sample_2014_refpoint), by = "pol_id") %>%
  mutate(distance = round(sqrt((x_refpoint - x_centroid)^2 + (y_refpoint - y_centroid)^2)))


```

We do not always get the same centroid as originally was obtained. 
In the table below we show the cases for which the centroid differs from the originally determined centroid. 
In three cases, this also results in a different grts-ranking.

```{r}
check_centroid %>%
  filter(distance > 0.1) %>%
  arrange(desc(distance)) %>%
  select(pol_id, "distance (m)" = distance, grts_ranking_original = grts_ranking, grts_ranking_centroid) %>%
  kable() %>%
  kable_styling()
```






Below an example for polygon wit pol_id = 19144.
The red is the originally calculated centroid, the blue dot the currently derived centroid.

```{r}

show_polygons <- sample_2014_sf %>%
  filter(pol_id == "19114") 

show_centroids <- sample_2014_ranking_centroid %>%
  filter(pol_id == "19114") 

show_refpoints <- sample_2014_refpoint %>%
  filter(pol_id == "19114") 

show_polygons %>%
ggplot() +
  geom_sf() +
  geom_sf(data = show_centroids,
          colour = "blue", size = 4) +
  geom_sf(data = show_refpoints,
          colour = "red", size = 4)
 

```

We will keep the originally determined centroids as reference points in case a different grts-ranking was obtained.
This way we can reproduce the orginally selected sample.

```{r}

sample_2014_refpoint <- check_centroid %>%
  mutate(x_coord = ifelse(grts_ranking != grts_ranking_centroid, x_refpoint, x_centroid),
         y_coord = ifelse(grts_ranking != grts_ranking_centroid, y_refpoint, y_centroid)) %>%
  st_as_sf(coords = c("x_coord", "y_coord"), crs = 31370) %>%
  select(pol_id, grts_ranking)

polygons_2014 <- sample_2014_sf %>%
  rename(grts_ranking_draw = grts_ranking) %>%
  left_join(st_drop_geometry(sample_2014_refpoint), by = "pol_id") %>%
  select(pol_id, sac, type_target_all, grts_ranking, grts_ranking_draw, set) 

sample_2014 <- sample_2014 %>%
  rename(grts_ranking_draw = grts_ranking) %>%
  left_join(st_drop_geometry(sample_2014_refpoint), by = "pol_id") %>%
  mutate(source_polygon = "habitatmap")

```



### Overview of number of sampling units per strata

```{r}
overzicht_sample_2014 <- sample_2014 %>%
  group_by(type_target, sac, area_class) %>%
  summarise(samplesize = n_distinct(pol_id)) %>%
  ungroup()
```
```{r}
overzicht_sample_2014 %>%
  spread(area_class, samplesize, fill = 0) %>%
  kable() %>%
  kable_styling()
```


## Available extra sampling units in 2016

In 2016 it appeared that the selected sampling units in 2014 were not sufficient to reach the desired sample size.
Additional sampling units were selected based on a preliminary version of the watersurfaces layer and the habitatmap version 2016.
Since then, a sampling unit corresponds to a polygon of the watersurfaces layer.
All watersurfaces that overlap with habitatmap polygons with 31xx habitat were selected, excluding habitatmap polygons in the 2014 sample that were already assessed in the field.
We drew also a sample for 2190_a by selecting all watersurfaces in habitatmap polygons that contain dune habitat, manually excluding some artificial watersurfaces that cleary dit not correspond with 2190_a. 

The files can be found in [this google drive folder](https://drive.google.com/drive/folders/1hXMgjTbiPWIgUv4cV3taNPRqAu7if1LS).

```{r}
sampleframe_31xx_2016_sf_orig <- read_sf("../data/watersurfaces_2016/Steekproef_31xx_extra_versie20160531.shp", crs = 31370)

sampleframe_31xx_2016_orig <- read_csv2("../data/watersurfaces_2016/Steekproef_31xx.csv")

sampleframe_2190_sf_orig <- read_sf("../data/watersurfaces_2016/Steekproefkader_2190_versie20170606.shp", crs = 31370) 

sampleframe_31xx_2016 <- sampleframe_31xx_2016_orig %>%
  select(pol_id = ID_Plas, type_target = Habsubt, sac = SBZH, area_class = OppKlasse, grts_ranking = Ranking, ranking_rel = RangNr, source_polygon = Type) %>%
  mutate(set = "selection2016",
         source_polygon = ifelse(source_polygon == "Polygoon blauwe laag", "watersurfaces", 
                                 ifelse(source_polygon == "Polygoon habitatkaart", "habitatmap", NA)))

polygon_id_type_target_all <- sampleframe_31xx_2016 %>%
  group_by(pol_id) %>%
  arrange(type_target) %>%
  summarise(type_target_all = str_c(type_target, collapse = ";")) %>%
  ungroup()

sampleframe_31xx_2016_sf <- sampleframe_31xx_2016_sf_orig %>%
  select(pol_id = ID_Plas,  grts_ranking = Ranking, sac = SBZH) %>%
  mutate(sac = as.numeric(sac)) %>%
  left_join(polygon_id_type_target_all, by = "pol_id") %>%
  mutate(set = "selection2016")

sampleframe_2190 <- sampleframe_2190_sf_orig %>%
  st_drop_geometry() %>%
  mutate(type_target = "2190") %>%
  select(pol_id = ID_plas, type_target, sac = SBZH, area_class = OppKlasse, grts_ranking = Ranking, ranking_rel = Volgorde) %>%
  mutate(set = "selection2016") %>%
  mutate(pol_id = as.character(pol_id),
         area_class = ifelse(area_class != "1 ha < Opp <= 5 ha", "Opp <= 1 ha", area_class),
         source_polygon = "watersurfaces")

sampleframe_2190_sf <- sampleframe_2190_sf_orig %>%
  mutate(type_target_all = "2190") %>%
  select(pol_id = ID_plas, grts_ranking = Ranking, sac = SBZH, type_target_all, selected = Steekproef) %>%
  mutate(pol_id = as.character(pol_id),
         set = "selection2016")
  
```

```{r}
refpoints_watersurfaces <- read_sf("../data/watersurfaces_2016/BlauweLaag_31xx_punten.shp", crs = 31370) %>%
  mutate(pol_id = str_c("BL_", ID_plas)) %>%
  select(pol_id) 

refpoints_habmap <- read_sf("../data/watersurfaces_2016/BWK_plassen_punten_20160524.shp", crs = 31370) %>%
  mutate(pol_id = str_c("BWK_", Pol_ID)) %>%
  select(pol_id)
     
refpoints_31xx <- refpoints_watersurfaces %>%
  bind_rows(refpoints_habmap) %>%
  inner_join(distinct( sampleframe_31xx_2016, pol_id, grts_ranking ), by = "pol_id")

refpoints_31xx <- refpoints_31xx %>%
  mutate(x_refpoint = st_coordinates(refpoints_31xx)[,1],
         y_refpoint = st_coordinates(refpoints_31xx)[,2])

```


### Reproduce grts_ranking based on original reference points

```{r}
refpoints_31xx_check <- refpoints_31xx %>%
  mutate(grts_ranking_check = grts_master[as(., "Spatial")]) 

```

The table below shows the 31xx polygons with a wrongly assigned grts-ranking based on the original reference points.
We will correct the grts-ranking use the variable grts_ranking_draw to store the wrongly assigned grts-ranking which was used to draw the sample.

```{r}
refpoints_31xx_check %>%
  st_drop_geometry() %>%
  filter(grts_ranking != grts_ranking_check) %>%
  kable() %>%
  kable_styling()
```

```{r}
refpoints_31xx <- refpoints_31xx_check %>%
  rename(grts_ranking_draw = grts_ranking, grts_ranking = grts_ranking_check)

```




### Reproduce referene points using point_on_surface

We used the point_on_surface function to select reference points for the sampling units.
We will check if we can reproduce the reference points for 31xx.

```{r}

check_refpoint_31xx <- sampleframe_31xx_2016_sf %>%
  st_point_on_surface() %>%
  mutate(grts_ranking_refpoint = grts_master[as(., "Spatial")]) %>%
  select(pol_id, grts_ranking_refpoint)

check_refpoint_31xx <- check_refpoint_31xx %>%
  mutate(x_refpoint_check = st_coordinates(check_refpoint_31xx)[,1],
         y_refpoint_check = st_coordinates(check_refpoint_31xx)[,2])

check_refpoint_31xx<- check_refpoint_31xx %>%
  st_drop_geometry() %>%
  left_join(st_drop_geometry(refpoints_31xx), by = "pol_id") %>%
  mutate(distance = round(sqrt((x_refpoint - x_refpoint_check)^2 + (y_refpoint - y_refpoint_check)^2)),
         grts_ranking_diff = grts_ranking_refpoint != grts_ranking)


```




For one 31xx polygon a different point was obtained using point_on_surface which also resulted in a different grts-ranking.
We will use the original reference point to be able to reproduce the sample.

```{r}
check_refpoint_31xx %>%
  filter(distance > 0) %>%
  select(pol_id, grts_ranking_refpoint, grts_ranking_orig = grts_ranking, distance) %>%
  kable() %>%
  kable_styling()


```






A wrong grts_ranking was obtained for 2190 sampling units (reading the shape file with grts-ranking values converted the grts-ranking values to factor levels)

We will use the variable grts_ranking_draw to indicate the originally used grts-ranking for drawing the sample.

```{r}

refpoint_2190 <- sampleframe_2190_sf %>%
  st_point_on_surface() %>%
  mutate(grts_ranking = grts_master[as(., "Spatial")]) %>%
  select(pol_id, grts_ranking)

refpoints_2016 <- refpoints_31xx %>%
  select(pol_id, grts_ranking) %>%
  bind_rows(refpoint_2190)

polygons_2016 <- sampleframe_31xx_2016_sf %>%
  bind_rows(sampleframe_2190_sf) %>%
  rename(grts_ranking_draw = grts_ranking) %>%
  left_join(st_drop_geometry(refpoints_2016), by = "pol_id") %>%
  select(pol_id, sac, type_target_all, grts_ranking, grts_ranking_draw, set) 

sampleframe_2016 <- sampleframe_31xx_2016%>%
  bind_rows(sampleframe_2190) %>%
  rename(grts_ranking_draw = grts_ranking) %>%
  left_join(st_drop_geometry(refpoints_2016), by = "pol_id") %>%
  mutate(area_class = ifelse(area_class == "> 50 ha", "Opp >= 50 ha", area_class),
         area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp < 50 ha", "Opp >= 50 ha"))) 

```

Below the selected sampling units for 2190_a (yellow polygons) and the other sampling units of the sampling frame (red polygons) that are not selected. Dispite the use of an incorrect grts-ranking, selected sampling units seem quite well distrubted at first sight.  

```{r}
sampleframe_2190_sf_orig %>%
  st_transform(4326) %>%
  mutate(color_show = ifelse(Steekproef == 1, "yellow", "red")) %>%
  leaflet() %>%
  addTiles() %>%
  addPolygons(color = ~color_show)
```
### Overview of number of sampling units per strata

```{r}
sampleframe_2016 %>%
  group_by(type_target, sac, area_class) %>%
  summarise(samplesize = n_distinct(pol_id)) %>%
  ungroup() %>%
  spread(area_class, samplesize, fill = 0) %>%
  kable() %>%
  kable_styling()
```

## Combine sample 2014 and sampleframe 2016

We combine the selected sampling units in 2014 and the sampling units of the sampleframe in 2016.

```{r}
mhq_watersurfaces_samplingunits <- polygons_2014 %>%
  bind_rows(polygons_2016) %>%
  mutate(sac = as.logical(sac))

mhq_watersurfaces_types <- sample_2014 %>%
  bind_rows(sampleframe_2016) %>%
  mutate(sac = as.logical(sac),
         area_class = ifelse(area_class == "> 50 ha", "Opp >= 50 ha", as.character(area_class)),
         area_class = factor(area_class, levels = c("Opp <= 1 ha", "1 ha < Opp <= 5 ha", "5 ha < Opp < 50 ha", "Opp >= 50 ha")))

mhq_watersurfaces_refpoints <- sample_2014_refpoint %>%
  bind_rows(refpoints_2016)


check_unique_pol_id <- nrow(mhq_watersurfaces_samplingunits) == n_distinct(mhq_watersurfaces_samplingunits$pol_id)
check_unique_ranking <- nrow(mhq_watersurfaces_samplingunits) == n_distinct(mhq_watersurfaces_samplingunits$grts_ranking)
check_types <- mhq_watersurfaces_types %>%
  group_by(pol_id, type_target) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n > 1)

watersurfaces_ranking_not_unique <- mhq_watersurfaces_samplingunits %>%
  group_by(grts_ranking) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n > 1)

```




### Check for sampling units without unique grts_ranking

Not all sampling units have a unique grts-ranking. See map below.

```{r}
watersurfaces_ranking_not_unique %>%
  st_transform(4326) %>%
  mutate(show_color = ifelse(set == "selection2014", "yellow", "red")) %>%
  leaflet() %>%
  addTiles() %>%
  addPolygons(label = ~grts_ranking, color = ~show_color)
```





Some sampling units selected in 2014 overlap with those of the 2016 sample frame. 


```{r}

sample_2014_overlap <- sample_2014_sf %>%
  st_join(select(sampleframe_31xx_2016_sf, pol_id_overlap = pol_id), largest = TRUE) %>%
  filter(!is.na(pol_id_overlap))

sampleframe_2016_overlap <- sampleframe_31xx_2016_sf %>%
  st_join(select(sample_2014_overlap, pol_id_overlap = pol_id), largest = TRUE) %>%
  filter(!is.na(pol_id_overlap))

sample_2014_overlap %>%
  st_transform(4326) %>%
  leaflet() %>%
  addTiles() %>%
  addPolygons(color = "yellow") %>%
  addPolygons(data = st_transform(sampleframe_2016_overlap, 4326), color = "red")
```




```{r, eval = TRUE}
mhq_watersurfaces_samplingunits %>%
  st_write(dsn = "../output/mhq_watersurfaces_original.gpkg", layer = "samplingunits_original", delete_layer = TRUE)

mhq_watersurfaces_refpoints %>%
  st_write(dsn = "../output/mhq_watersurfaces_original.gpkg", layer = "refpoints_original", delete_layer = TRUE)


```


