# Update sampling frame

We will update the sampling frame based on the `watersurfaces_hab` (version v4) data source.


```{r}
watersurfaces_hab <- read_watersurfaces_hab()

path <- fileman_up("n2khab_data")
file <- "20_processed/watersurfaces_hab/watersurfaces_hab.gpkg"

md5 <- file.path(path, file) %>%
  file() %>%
  md5() %>%
  str_c(collapse = '')

md5_ref <- "96b6a7abc3d637d71052900970e70904"

tibble(
  file = file.path(path, file), md5, md5_ref
) %>%
  kable() %>%
  kable_styling()

watersurfaces_hab_types <- watersurfaces_hab$watersurfaces_types %>%
  filter(! str_detect(type, "rbb")) %>%
  mutate(main_type = str_sub(type, 1, 4)) %>%
  group_by(polygon_id, main_type) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(!((type == main_type) & (n > 1))) %>%
  select(-n, -main_type) #bij sommige polygonen zowel 3130 (meestal 3130_rbb..) als 3130_aom als type --> enkel 3130_aom behouden 

```

## Updating watersurface reference points


The reference point of a watersurface is the point location based on which the grts_ranking of the watersurface is defined. 

We update the table of watersurface reference points as follows:

+ We start from the table of reference points (`watersurfaces_refpoints`) that is based on the original design of the mhq schemes and the assessments until 2020.
+ Next, we identify watersurfaces from watersurface_hab that are not in the original watersurfaces_refpoints table and define the refpoint using `point_on_surface`. 
+ Then we assign the grts_ranking to the reference point based on the `GRTSmh` data source 
+ finally we add the new reference points to `watersurfaces_refpoints` 

```{r}
grts_master <- read_GRTSmh()

watersurfaces_refpoints <- read_vc(file = "watersurfaces_refpoints", root = "../output")

refpoints_new <- watersurfaces_hab$watersurfaces_polygons %>%
  anti_join(watersurfaces_refpoints, by = "polygon_id") %>%
  # semi_join(mhq_watersurfaces_samplingunits_new, by = "code_watersurfaces") %>%
  st_point_on_surface() %>%
  mutate(grts_ranking = grts_master[as(., "Spatial")],
         in_object = TRUE)

refpoints_new <- refpoints_new %>%
  st_drop_geometry() %>%
  mutate(x = round(st_coordinates(refpoints_new)[,1], 2),
         y = round(st_coordinates(refpoints_new)[,2], 2)) %>%
  select(polygon_id, grts_ranking, in_object, x, y)
    
watersurfaces_refpoints_update <- watersurfaces_refpoints %>%
  bind_rows(refpoints_new) %>%
  select(polygon_id, everything())

check <- watersurfaces_refpoints_update %>%
  group_by(polygon_id) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n > 1)

watersurfaces_refpoints_update %>%
  write_vc("watersurfaces_refpoints", "../output", sorting = c("polygon_id"))

```


```{r}
watersurfaces_refpoints_update %>%
  slice(1:5)
```

## Updating mhq watersurfaces population units

The table `mhq_watersurfaces_populationunits` contains all population units in the sample frame of the mhq schemes.
The original table is based on the original design of mhq.
We want to keep the grts_ranking and grts_ranking_draw of the original design to assure maximum overlap between samples drawn from different versions of the sample frame.

We update the table as follows:

+ From the `watersurface_hab` data source we select all watersurface/type combinations that are not in the original sample frame.
+ We add the grts_ranking to the selected watersurfaces from the `watersurfaces_refpoints` table
+ We add the new sampling units to the `mhq_watersurfaces_populationunits` table. 
+ Based on the mhq assessments we check if the target type was observed in the assessed watersurfaces. If not the watersurface is removed from the table. 
+ Population units (polygon_id/type combination) from the original sampling frame that not have been assessed yet and that do not occur in the latest version of watersurfaces_hab are removed 

```{r}

mhq_watersurfaces_populationunits <- read_vc(file = "mhq_watersurfaces_populationunits", root = "../output")

mhq_watersurfaces_populationunits_3130 <- mhq_watersurfaces_populationunits %>%
  filter(str_sub(type, 1, 4) == "3130")

mhq_watersurfaces_populationunits_new <-  watersurfaces_hab_types %>%
  distinct(polygon_id, type) %>%
  anti_join(mhq_watersurfaces_populationunits, by = c("polygon_id", "type")) %>%
  filter(!(polygon_id %in% mhq_watersurfaces_populationunits_3130$polygon_id &
            type == "3130")) %>%
  # filter(!str_detect(polygon_id, "_v")) %>%
  mutate(source = "watersurfaces_hab_v4") %>%
  left_join(select(watersurfaces_refpoints_update, polygon_id, grts_ranking), by = "polygon_id") %>%
  mutate(scheme = ifelse(type == "2190_a", "HQ2190_aq", str_c("HQ", str_sub(type, 1, 4))),
         scheme = factor(scheme, levels = levels(schemes_mhq$scheme)),
         grts_ranking_draw = grts_ranking,
         sampling_unit_code = str_c(polygon_id, "_", type)) 

mhq_watersurfaces_populationunits_update <- mhq_watersurfaces_populationunits %>%
  bind_rows(mhq_watersurfaces_populationunits_new)
  
mhq_watersurfaces_assessments <- read_vc(file = "mhq_watersurfaces_assessments", root = "../output")

watersurfaces_populationunits <- mhq_watersurfaces_populationunits_update %>%
  left_join(mhq_watersurfaces_assessments, by = c("polygon_id", "type")) %>%
  filter(is_present | is.na(is_present)) %>%
  select(scheme, sampling_unit_code, polygon_id, type, grts_ranking, grts_ranking_draw, source)

# watersurfaces_hab_3130 <- watersurfaces_hab_types %>%
#   filter(type == "3130")
# 
# watersurfaces_populationunits_remove <- watersurfaces_populationunits %>%
#   filter(!assessed) %>%
#   anti_join(watersurfaces_hab_types, by = c("polygon_id", "type")) %>%
#   filter(! (str_sub(type, 1,4) == "3130" & polygon_id %in% watersurfaces_hab_3130$polygon_id))
# 
# watersurfaces_populationunits <- watersurfaces_populationunits %>%
#   anti_join(watersurfaces_populationunits_remove, by = c("polygon_id", "type")) %>%
#   rename(type = type) %>%
#   unique()

check <- watersurfaces_populationunits %>%
  group_by(polygon_id, type) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n > 1)

remove <- watersurfaces_populationunits %>%
  filter(source == "original sample frame") %>%
  anti_join(watersurfaces_hab_types, by = c("polygon_id", "type"))

gh <- mhq_watersurfaces_assessments %>%
  filter(!any_habitat)

watersurfaces_populationunits_select <- watersurfaces_populationunits %>%
  anti_join(remove, by = c("polygon_id", "type")) %>%
  anti_join(gh, by = "polygon_id")

check <- watersurfaces_populationunits_select %>%
  group_by(polygon_id, type) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n > 1)

watersurfaces_populationunits_select %>%
  write_vc("mhq_watersurfaces_populationunits", root = "../output", sorting = c("type", "grts_ranking_draw", "polygon_id"), strict = FALSE)

```

```{r}
watersurfaces_populationunits_select %>%
  slice(1:5, 400:405)
```



