# Updating 3260 sampling frame

We will update the old sampling frame for 3260 (that was based on 3260 map version 1.4). 
The old sampling frame consists of points that represent the sampling units (100 meter river segments). 
Each point is located in the most downstream part of the sampling unit. 
We will start with selecting points each 100 meter for all rivers in Flanders so that when the 3260 map changes in the future, we do not need to select new points again.
Of course we want to keep the the points in the old sample frame as quite a number of them have already been sampled and measured in the field. 
So we will exclude the newly selected points that overlap with the old 3260 map and replace this with the points in the old sample frame.
Finally we will select the points that overlay with the new version of the 3260 map resulting in the updated sampling frame.
The stepwise procedure is explained below. 

## Data import

- Read new 3260 map (v1.7) and sample frame based on old 3260 map. See this [Google Drive folder](https://drive.google.com/drive/folders/1mZJH1vaED6s2BA-zQivn8Nrag9NZ4C9D).

```{r}

map_3260_new <- read_sf("../data/habitatstreams/v1_7", "BasisbestandSteekproefvernieuwingHT3260_SBZ", crs = 31370)

map_3260_old <- read_sf("../data/habitatstreams/v1_4", "BasisbestandSteekproeftrekkingHT3260", crs = 31370)

segments_3260_old <- read_vc(file = "sample_habitatstreams_v1_4", "../output")

segments_3260_old_sf <- segments_3260_old %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) 
```


- Read Flemish Hydrological Atlas (VHA) with all river segments in Flanders. We use the version of VHA that was used to create habitatstreams v1.7: 201806. See this [Google Drive folder](https://drive.google.com/drive/folders/1ikEkf5N4ckWcyCMSgM1O_OgcZCFjX9B4).  

```{r}
vha <- read_sf("../data/VHA_201806", "Wlas", crs = 31370)
```


- Read sac

```{r}
sac <- read_admin_areas(dsn = "sac") %>%
  dplyr::select(sac_code)
```

- Read grts

```{r}
grts_master <- read_GRTSmh()
```



## Select points that represent 100m sampling units on all rivers in Flanders

- select points on line segments in vha every 100 meters

```{r}
vha_length <- sum(vha$LENGTE)

segments_vha <- spsample(as_Spatial(vha),
                          n = ceiling(vha_length/100),
                          type="regular",
                          cellsize=100,
                          offset=0)

vha_buffer <- st_buffer(vha, dist = 1)
                              
segments_vha$name <- over(segments_vha, as_Spatial(vha_buffer))$NAAM
segments_vha$wl_code <- over(segments_vha, as_Spatial(vha_buffer))$WTRLICHC
segments_vha$vhas <- over(segments_vha, as_Spatial(vha_buffer))$VHAS
segments_vha$oidn <- over(segments_vha, as_Spatial(vha_buffer))$OIDN

segments_vha <- segments_vha %>%
  as("sf")

st_crs(segments_vha) <- st_crs(vha)

#st_write(segments_vha, "../output/segments_vha", "segments_vha", driver = "ESRI Shapefile")
```


- Next we select the points that are not within the old map of 3260 (version 1.4), because for this old map we have already selected points every 100 meter and we want to keep the previously selected points.

- We also want to avoid that the newly selected point on the vha linestrings are closer than 100 meter to the previously selected points when both are in the same waterbody. 
We take a (linear) buffer of 50 meter along the previously selected points. So to avoid to much complexity, we do not calculate distances along the river linestrings. 
We leave out the newly selected points that are within a buffer of a previuosly selected point that is located within the same waterbody (defined bij wl_code).

- Note that the wl_code assigned to the previously selected points are outdated and have to be updated based on an overlay with vha.

```{r}

vha_buffer <- st_buffer(vha, dist = 1) %>%
  dplyr::select(wl_code = WTRLICHC)

segments_3260_old_sf <- segments_3260_old %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  rename(wl_code_old = wl_code) %>%
  st_join(vha_buffer, join = st_nearest_feature)
    
segments_3260_old_buffer <- segments_3260_old_sf %>%
  st_buffer(dist = 50) %>%
  dplyr::select(segment_id, wl_code_buffer = wl_code)

#st_write(segments_3260_old_buffer, "../output/check.gpkg", "segments_3260_old_buffer", driver = "GPKG")

map_3260_old_buffer <- map_3260_old %>%
  st_buffer(dist = 1) %>%
  dplyr::select(wl_code_old_3260 = WLcode)

vha_overlay <- segments_vha %>%
  mutate(id = str_c("new_", 1:nrow(.))) %>%
  st_join(map_3260_old_buffer) %>%
  mutate(in_map_3260_old = !is.na(wl_code_old_3260),
         x = st_coordinates(.)[,1],
         y = st_coordinates(.)[,2],
         grts_ranking = grts_master[as(., "Spatial")]) %>%
  # group_by(id) %>%
  # mutate(dist_to_segments_old = min(drop_units(st_distance(geometry, segments_3260_old_sf)))) %>%
  # ungroup() %>%
  st_join(sac) %>%
  st_join(segments_3260_old_buffer) 
  
#st_write(vha_overlay, "../output/vha_overlay", "vha_overlay", driver = "ESRI Shapefile")

vha_overlay_df <- vha_overlay %>%
  st_drop_geometry()

vha_difference <- vha_overlay_df %>%
  filter((!in_map_3260_old) | (in_map_3260_old & is.na(wl_code_buffer)))  %>%
  mutate(check = wl_code != wl_code_buffer)


vha_difference_buffer <- vha_difference %>%
  group_by(id, name, wl_code, sac_code, grts_ranking, x, y) %>%
  summarise(within_100mbuffer_3260_old = sum(wl_code == wl_code_buffer, na.rm = TRUE) > 0) %>%
  ungroup() %>%
  filter(!within_100mbuffer_3260_old) 



```


- Next we combine the newly selected vha points with the previously selected point on the old 3260 map. We also add the strata information (position relative to sac) that is part of the old 3260 map.

```{r}
map_3260_old_strata <- map_3260_old %>%
  mutate(sac_old = ifelse(SBZ == "sa", "downstream",
                          ifelse(SBZ == "intersect SBZ", "intersect",
                                 ifelse(SBZ == "buiten", "outside", NA)))) %>%
  dplyr::select(sac_old, naam) %>%
  st_buffer(dist = 1) 
  # st_join(sac, largest = TRUE) %>%
  # rename(sac_code_segment = sac_code)


vha_segments_combined <- segments_3260_old_sf %>%
  st_join(sac) %>%
  st_join(map_3260_old_strata, largest = TRUE) %>%
  mutate(id = str_c("old_", segment_id),
         x = st_coordinates(.)[,1],
         y = st_coordinates(.)[,2]) %>%
  st_drop_geometry() %>%
  dplyr::select(id, name = river_name, wl_code, wl_code_old, sac_code, sac_old, grts_ranking, x, y) %>%
  bind_rows(vha_difference_buffer)
  
#vha_segments_combined %>%
  # st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  # st_write(dsn = "../output/check2.gpkg", "vha_segments_combined", driver ="GPKG")
```


## Create updated sample frame for 3260

We create an updated sample frame for 3260 by selecting the points that intersect with a buffer of 5 meter around the new 3260 habitatmap (v1.7). 

We also want to include points that do not intersect but for which the 100 m segment will contain 3260 habitat according to the habitatmap. 
Remember that the points represent the most downstream location of each 100 m segment. 
So we only want to select (not intersecting points) that are located (less than 100 meter) downstream of 3260 habitat. 
We do this by creating a buffer of 75 meter around the starting points of each linestring in the habitatmap (these starting points are normally the most downstream location of the linesegment) and than selecting all points that intersect with the buffer and have the same wl_code (to avoid selectin points in other streams). 

```{r}

map_3260_new_buffer <- map_3260_new %>%
  st_buffer(10) %>%
  st_combine() %>%
  st_geometry()

#st_write(map_3260_new_buffer, "../output/check.gpkg", "map_3260_new_buffer", driver = "GPKG")

map_3260_new_sac <- map_3260_new %>%
  st_buffer(10) %>%
  dplyr::select(SBZ)

segments_3260_new_intersect <- vha_segments_combined %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  #mutate(check = st_intersects(geometry, map_3260_new_buffer, sparse = FALSE)) %>%
  # st_join(map_3260_new_buffer, largest = TRUE)
  # st_drop_geometry()
  # filter(st_intersects(geometry, map_3260_new_buffer, sparse = FALSE)) %>%
  st_join(map_3260_new_sac, largest = TRUE) %>%
  filter(!is.na(SBZ))

map_3260_new_startpoint_buffer <- st_startpoint(map_3260_new) %>%
  st_as_sf() %>%
  mutate(id_temp = 1:nrow(.)) %>%
  st_join(map_3260_new) %>%
  st_join(vha_buffer) %>%
  st_buffer(dist = 75) %>%
  dplyr::select(SBZ, wl_code_extra = wl_code) %>%
  group_by(SBZ, wl_code_extra) %>%
  summarise(n = n()) %>%
  ungroup()

segments_3260_new_extra <- vha_segments_combined %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  st_join(map_3260_new_startpoint_buffer, largest = TRUE) %>%
  filter(!is.na(SBZ)) %>%
  filter(wl_code == wl_code_extra) %>%
  filter(!(id %in% segments_3260_new_intersect$id))

check <- segments_3260_new_extra %>%
  st_drop_geometry() %>%
  group_by(id) %>%
  summarise(n = n()) %>%
  ungroup()

segments_3260_new <- segments_3260_new_extra %>%
  dplyr::select(-wl_code_extra, -n) %>%
  rbind(segments_3260_new_intersect) %>%
  mutate(sac = ifelse(SBZ %in% c("sa", "intersect SBZ"), "intersecting or downstream",
                      ifelse(SBZ %in% c("buiten"), "outside", NA)),
         catchment_area = ifelse(str_sub(wl_code, 1, 2) %in% c("VL", "L1"), "> 10 km²",
                                  ifelse(str_sub(wl_code, 1, 2) == "L2",  "< 10 km²", "< 10 km²"))) %>%
  dplyr::select(segment_code = id, river_name = name, wl_code, wl_code_old, grts_ranking, sac, catchment_area)
```


```{r}

segments_3260_border <- segments_3260_new %>%
  filter(is.na(grts_ranking))

segments_3260_border_buffer <- segments_3260_border %>%
  st_buffer(dist = 100)

grts_border <- grts_master %>%
  mask(segments_3260_border_buffer) %>%
  rasterToPoints(spatial = TRUE) %>%
  st_as_sf()

st_crs(grts_border) <- 31370

#write_sf(grts_border, dsn = "../output/grts_border4", layer = "check_grts_border", driver = "ESRI Shapefile")

segments_3260_border_grts <- segments_3260_border %>%
 st_join(grts_border, join = st_nearest_feature) %>%
  group_by(segment_code) %>%
  mutate(dist = min(drop_units(st_distance(geometry, grts_border)))) %>%
  ungroup() %>%
  mutate(grts_ranking = GRTSmaster_habitats) %>%
  # filter(dist < 32 | name == "MAAS") %>%
  dplyr::select(-GRTSmaster_habitats, -dist) 

segments_3260_new_addgrts <- segments_3260_new %>%
  filter(!is.na(grts_ranking)) %>%
  rbind(segments_3260_border_grts) 

sample_frame_3260 <- segments_3260_new_addgrts %>%
  mutate(x = st_coordinates(geometry)[,1],
         y = st_coordinates(geometry)[, 2]) %>%
  unique() %>%
  rename(segment_id = segment_code) %>%
  mutate(segment_id = str_replace(segment_id, "old", "1_4"),
         segment_id = str_replace(segment_id, "new", "1_7")) %>%
  dplyr::select(segment_id, grts_ranking, sac_streams = sac, x, y)


check <- sample_frame_3260 %>%
  st_drop_geometry() %>%
  group_by(segment_id) %>%
  mutate(n = n()) %>%
  ungroup()

# correction based on visual inspection
sample_frame_3260 <- sample_frame_3260 %>%
  filter(!(segment_id == "1_4_1873" & sac_streams == "outside")) %>%
  filter(!(segment_id == "1_7_17866" & sac_streams == "outside"))
  

write_vc(st_drop_geometry(sample_frame_3260), root =  "../output", file = "samplingframe_habitatstreams_v1_7", sorting = "segment_id", strict = FALSE)

st_write(sample_frame_3260, "../output/samplingframe_streams.gpkg", "samplingframe_habitatstreams_v1_7", driver = "GPKG", append = TRUE, delete_layer = TRUE)

streams_segments_all <- vha_segments_combined %>%
  mutate(segment_id = str_replace(id, "old", "1_4"),
        segment_id = str_replace(segment_id, "new", "1_7") ,
        name = str_to_lower(name)) %>%
  dplyr::select(segment_id, river_name = name, wl_code, x, y, grts_ranking) 

check_unique_id <- streams_segments_all %>%
  group_by(segment_id) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(n > 1)

streams_segments_all %>%
  write_vc(root =  "../output", file = "vha_segments", sorting = c("segment_id", "grts_ranking", "river_name"), strict = FALSE)

```









