# Update sample frame

## New 3260 map

The sample frame is based on the 3260 habitat type map (version [habitatstreams_2023](https://zenodo.org/records/10353508)).

In this map following categories were (manually) determined:

+ intersecting or downstreams of sac
+ other

The result can be found in the file 'BasisbestandSteekproefvernieuwingHT3260_SBZ_2025.shp', which is stored in [this google drive](https://drive.google.com/drive/folders/1hyhgpxJWa9mSWttgM_l1bDww2r0uEx6-) folder 

```{r}
map3260 <- st_read("../data/map3260_sbzh_2025/BasisbestandSteekproefvernieuwingHT3260_SBZ_2025.shp")

colnames(map3260) <- str_to_lower(colnames(map3260) )

map3260 <- map3260 %>%
  mutate(sac_streams = ifelse(sbz %in% c("intersect SBZ", "sa"), "intersecting or downstream", 
                              ifelse(sbz == "buiten", "outside", NA)))
```

## Watercourse 100mseg

The data source [watercourse_100mseg](https://zenodo.org/records/4452578) represents all officially known watercourses of the Flemish Region as line segments of 100 m.
This data source can be used to select sampling units.

```{r}
watercourse_100mseg <- read_watercourse_100mseg()

watercourse_100mseg_points <- watercourse_100mseg$points
  
watercourse_100mseg_lines <- watercourse_100mseg$lines
```

## River segments with 3260 habitat

We want to select all 100 m river segments that contain 3260 habitat.

Workflow:

+ 1 meter buffer around 3260 map
+ select all watercourse segments that intersect with 3260 (including buffer)
+ calculate the length of each segment within the 1 meter buffer 3260 map
+ remove segments with length <= 2 meter (these are segments that are branches from a 3260 linestring)

```{r}

map3260_buffer1 <- map3260 %>%
  st_buffer(1)

watercourse_100mseg_3260 <- watercourse_100mseg_lines %>%
  st_join(map3260_buffer1, largest = TRUE) %>%
  filter(!is.na(objectid))

watercourse_100mseg_3260intersection <- watercourse_100mseg_lines %>%
  st_intersection(map3260_buffer1) %>%
  mutate(length_3260 = st_length(geom))

watercourse_100mseg_prop3260 <- watercourse_100mseg_3260intersection %>%
  st_drop_geometry() %>%
  group_by(rank, vhag_code) %>%
  summarise(prop_3260 = round(sum(drop_units(length_3260)), 0)) %>%
  ungroup() %>%
  filter(prop_3260 > 2)

watercourse_100mseg_3260 <- watercourse_100mseg_3260 %>%
  inner_join(watercourse_100mseg_prop3260, by = c("rank", "vhag_code"))

```

## Check result 

+ Are all objects in the 3260 map covered by a watercourse segment?

Following segments are not covered

```{r}

map3260_watercourse <- map3260_buffer1 %>%
  st_join(select(watercourse_100mseg_3260, rank, prop_3260))

map3260_watercourse %>%
  filter(is.na(rank)) %>%
  st_drop_geometry() %>%
  kable()
```

One example: Oude Molebeek.

```{r}
map3260_object <- map3260 %>%
  filter(objectid == 276)

watercourse_100mseg_select <- watercourse_100mseg_lines %>%
  st_join(st_buffer(map3260_object, 100)) %>%
  filter(!is.na(objectid))

map3260_object %>%
  st_transform(4326) %>%
  leaflet() %>%
  addTiles() %>%
  addPolylines() %>%
  addPolylines(data = st_transform(watercourse_100mseg_select, 4326), color = "black")
```

+ do we get the expected number of overlapping watercourse segments per 3260 line segment?

```{r}
length_3260 <- map3260 %>%
  mutate(length_m = round(drop_units(st_length(geometry)),2),
         n_expected = ceiling(length_m / 100)) %>%
  st_drop_geometry() %>%
  select(objectid,length_m, n_expected)

check_overlap <- map3260_watercourse %>%
  filter(!is.na(rank)) %>%
  st_drop_geometry() %>%
  group_by(objectid) %>%
  summarise(n_overlap = n_distinct(rank)) %>%
  ungroup() %>%
  left_join(length_3260, by = "objectid") %>%
  mutate(diff_n = abs(n_expected - n_overlap)) %>%
  filter(diff_n > 1)

check_overlap %>%
  kable()
```



An example where we get more overlapping segments then expected:

```{r}
map3260_object <- map3260_watercourse %>%
  filter(objectid == 42)

watercourse_100mseg_select <- watercourse_100mseg_3260 %>%
  filter(rank %in% map3260_object$rank)

map3260_object %>%
  st_transform(4326) %>%
  leaflet() %>%
  addTiles() %>%
  addPolylines() %>%
  addPolylines(data = st_transform(watercourse_100mseg_select, 4326), color = "black", label = ~prop_3260)
```

This is OK.
The 2 extra overlapping segments are covered by other 3260 line segments (resulting in a large prop_3260).

An example where we get less overlapping segments then expected:

```{r}
map3260_object <- map3260_watercourse %>%
  filter(objectid == 13)

watercourse_100mseg_select <- watercourse_100mseg_3260 %>%
  filter(rank %in% map3260_object$rank)

map3260_object %>%
  st_transform(4326) %>%
  leaflet() %>%
  addTiles() %>%
  addPolylines() %>%
  addPolylines(data = st_transform(watercourse_100mseg_select, 4326), color = "black")
```

This is not OK but only occurs in two occasions.

## Overall overlap

```{r}
watercourse_100mseg_3260_lines <- watercourse_100mseg_3260 %>%
  select(sac_streams, naam, rank, vhag_code,  watercourse_code = wtrlichc, prop_3260)

length_segments <- watercourse_100mseg_3260_lines %>%
  mutate(length = st_length(geom)) %>%
  st_drop_geometry() %>%
  summarise(length = sum(drop_units(length) * prop_3260/ 100 / 1000))

length_map <- map3260 %>%
  mutate(length = st_length(geometry)) %>%
  st_drop_geometry() %>%
  summarise(length = sum(drop_units(length) / 1000))

overlap <- length_segments/length_map
```

The total length of 3260 habitat in all the selected watercource segments is : `r round(length_segments$length)` km.

The total length of 3260 habitat in the habitat map is : `r round(length_map$length)` km.

The overlap is `r round(overlap * 100, 1)` %.

## Write result

```{r}

check_unique_rank <- (watercourse_100mseg_3260_lines %>%
  st_drop_geometry() %>%
  distinct(rank) %>%
  nrow()) ==
  (watercourse_100mseg_3260_lines %>%
  st_drop_geometry()) %>%
  nrow()
    
watercourse_100mseg_3260_points <- watercourse_100mseg_points %>%
  filter(rank %in% watercourse_100mseg_3260_lines$rank)

watercourse_100mseg_3260_lines %>%
  st_write(dsn = "../output/watercourse_100mseg_3260.gpkg", "watercourse_100mseg_3260_lines", delete_layer = TRUE)

watercourse_100mseg_3260_points %>%
  st_write(dsn = "../output/watercourse_100mseg_3260.gpkg", "watercourse_100mseg_3260_points", delete_layer = TRUE)
```

