# Progress cycle 2

The second monitoring cycle of the 3260 mhq scheme started in 2022 and last until 2027.

It appears that many of the selected sampling units do not contain 3260 habitat. Therefore we have to select new sampling units to achieve a sufficient sample size.

## Selected sampling units

The file 'meetpunten3260_cyclus2_versie2022-05-05.csv' list the selected sampling units for cycle 2.
The file can be found in this [google drive folder](https://drive.google.com/drive/folders/1UNKCLEgiYq2r6EkfWXMkppSDraPcqfAU).

```{r}
sample_cycle2 <- read_csv2("../data/meetpunten3260_cyclus2_versie2022-05-05.csv")
```
```{r}
sample_cycle2 %>%
  group_by(sac_streams) %>%
  summarize(sample_size = n()) %>%
  ungroup() %>%
  kable() %>%
  kable_styling()
```



## Administration

The file 'Monitoring3260_SVZ2024_voorToonWestra.csv' contains the progress of the field work.
It can be found in this [google drive folder](https://drive.google.com/drive/folders/1DHLwdrp7i_rIwIfrj3-Tm_yXhvdm19D4).

```{r}
admin_orig <- read_csv2("../data/Monitoring3260_SVZ2024_voorToonWestra.csv")

admin <- admin_orig %>%
  mutate(assessment_year = `jaar monitoring`,
         assessment_date = as.Date(`datum veldbezoek`, format = "%d/%m/%Y"),
         sac_streams = ifelse(SBZ == 1, "intersecting or downstream",
                      ifelse(SBZ == 0, "outside", SBZ)),
         type_observed = doelhabitat) %>%
  select(field_code = Veldcode, sampling_unit_code, assessment_year, assessment_date, type_observed, klasse, reason_no_assessment = `reden geen bemonstering/ongeschikt`, x_assessment = x, y_assessment = y, position_change = `locatie SO of SA verschoven (xy aangepast tov origineel aangereikt meetpunt, volgens protocol)`) %>%
  filter(!is.na(field_code))

admin_cycle2 <- admin %>%
  filter(assessment_year >= 2022)


```

## Progress

```{r}
sample_cycle2_progress <- sample_cycle2 %>%
  select(-field_code) %>%
  left_join(admin_cycle2, by = "sampling_unit_code") %>%
  mutate(klasse = ifelse(is.na(klasse), "nog niet bemonsterd", klasse))

sample_cycle2_nt_bemonsterd <- sample_cycle2_progress %>%
  filter(klasse == "nog niet bemonsterd") %>%
  select(colnames(sample_cycle2))

sample_cycle2_nt_bemonsterd %>%
  write_csv2("../output/3260_cycle2_niet_bezocht.csv")
```

```{r}
progress_cycle2 <- sample_cycle2_progress %>%
  group_by(klasse, reason_no_assessment) %>%
  summarise(n_sampling_units = n()) %>%
  ungroup() 

progress_cycle2 %>%
  kable() %>%
  kable_styling()
```

```{r}
progress_sample_size <- sample_cycle2_progress %>%
  mutate(status_su = ifelse(!is.na(reason_no_assessment), 
                            ifelse(klasse == "niet meetellen (is vervangen)", "replaced", "not_measured"),
                            ifelse(klasse == "nog niet bemonsterd", "to_do", "measured")
                          )) %>%
  group_by(sac_streams, status_su) %>%
  summarise(n = n()) %>%
  ungroup()

```

