# Add GRTS ranking

## Add GRTS ranking to 3260 watercourse segment points

We assign a GRTS ranking to every 3260 watercourse segment point based on the [GRTS master covering Flanders](https://zenodo.org/records/2682323).

The GRTS master does not contain grid cells with a centroid outside Flanders.
Therefore some watercourse segment points near the border are not located in a GRTS master grid cell.
In that case, we select the GRTS ranking of the closest grid cell. 

```{r}
grts_master <- read_GRTSmh()

watercourse_100mseg_3260_ranking <- watercourse_100mseg_3260_points %>%
  mutate(grts_ranking = (terra::extract(grts_master, watercourse_100mseg_3260_points))$GRTSmaster_habitats) %>%
  left_join(watercourse_100mseg_3260_lines %>%
              st_drop_geometry() %>%
              select(rank, sac_streams, prop_3260),
            by = "rank")

no_ranking <- watercourse_100mseg_3260_ranking %>%
  filter(is.na(grts_ranking)) %>%
  select(-grts_ranking)

grts_border <- grts_master %>%
  mask(st_buffer(no_ranking, 100)) %>%
  as.points() %>%
  st_as_sf() %>%
  rename(grts_ranking = GRTSmaster_habitats)

no_ranking <- no_ranking %>%
  st_join(grts_border, join = st_nearest_feature) %>%
  group_by(geom) %>%
  mutate(dist = min(st_distance(geom, grts_border))) %>%
  ungroup()
```

The map below shows the watercourse segments with the most downstream point > 50 meters from a GRTS cell centroid.

```{r}

check_no_ranking <- no_ranking %>%
  filter(drop_units(dist) > 50)

check_lines <- watercourse_100mseg_3260_lines %>%
  filter(rank %in% check_no_ranking$rank) %>%
  st_transform(crs = 4326)

grts_border %>%
  filter(grts_ranking %in% check_no_ranking$grts_ranking) %>%
  st_transform(crs = 4326) %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(label =~ grts_ranking) %>%
  addCircleMarkers(data = st_transform(check_no_ranking, crs = 4326), label =~rank) %>%
  addPolylines(data = check_lines)
```


```{r}
watercourse_100mseg_3260_ranking <- watercourse_100mseg_3260_ranking %>%
  filter(!is.na(grts_ranking)) %>%
  bind_rows(no_ranking) %>%
  st_drop_geometry()

# some watercourse segments have the same ranking
check <- watercourse_100mseg_3260_ranking %>%
  group_by(grts_ranking) %>%
  filter(n() > 1)
```


## Add GRTS ranking draw based on original sample frame

The original sample frame `samplingframe_habitatstreams_v1_7` consists of points representing the most downstream point of 100 meter segment. 
It is based on [version 1.7 of the 3260 habitat map](https://zenodo.org/records/4420858).

These segment points do not overlap with the watercourse segment points of [watercourse_100mseg](https://zenodo.org/records/4452578) data source.

To reproduce the selection of sampling units in MHQ 3260 we will define a variable `grts_ranking_draw` for the 3260 `watercourse_100mseg` segments that overlap with the original sample frame.
`grts_ranking_draw` corresponds to the GRTS ranking of the best matching `samplingframe_habitatstreams_v1_7` segment.

The GRTS ranking of measured sampling units is assigned to the `watercourse_100mseg` segment with the largest overlap.

The GRTS ranking of other sampling units is assigned to the closest `watercourse_100mseg` segment point.

```{r}
# old sample frame

sample_frame_3260_v1_7_orig <- read_vc(root =  "../../habitatstreams/output", file = "samplingframe_habitatstreams_v1_7")

sample_frame_3260_v1_7 <- sample_frame_3260_v1_7_orig %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) 

```

```{r}

sample_cycle2_progress <- sample_cycle2_progress %>%
  rename(segment_id = sampling_unit_code)

sample_cycle2_keep <- sample_cycle2_progress %>%
  filter(!(klasse == "nog niet bemonsterd" & type_meetpunt == "Nieuw meetpunt")) %>%
  mutate(position_change = ifelse(is.na(position_change), "nee", position_change),
         x = ifelse(position_change == "ja" , x_assessment, x),
         y = ifelse(position_change == "ja", y_assessment, y)) %>%
  filter(field_code != "KaaBe_01") %>%
  select(segment_id, grts_ranking_draw, sac_streams, klasse, x, y) %>%
  mutate(cycle = 2)

```

```{r, eval =FALSE}
mhq_streams_refpoints_validity <- read_vc(root =  "../../habitatstreams/output", file = "mhq_streams_refpoints_validity_v2022")

mhq_streams_valid <- mhq_streams_refpoints_validity %>%
  filter(is_valid_refpoint)

mhq_streams_refpoints <- read_vc(root =  "../../habitatstreams/output", file = "mhq_streams_refpoints_v2022")

mhq_streams_refpoints_valid <- mhq_streams_refpoints %>%
  semi_join(mhq_streams_valid, by = "sampling_unit_code") 

mhq_streams_refpoints_keep <- mhq_streams_refpoints %>%
  semi_join(mhq_streams_valid, by = "sampling_unit_code") %>%
  anti_join(sample_cycle2_progress, by = "segment_id") %>%
  select(segment_id, grts_ranking_draw, sac_streams, x, y)%>%
  mutate(cycle = 1)

sample_unit_keep <- sample_cycle2_keep %>%
  bind_rows(mhq_streams_refpoints_keep) %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  group_by(geometry) %>%
  mutate(dist_to_3260map = round(min(st_distance(map3260, geometry)), 1 )) %>%
  ungroup() %>%
  filter(!segment_id == "1_7_27869") #foutief punt

``` 

```{r}
mhq_streams_assessments <- read_vc(root =  "../data/v2024", file = "mhq_streams_assessments")
mhq_streams_refpoints <- read_vc(root =  "../data/v2024", file = "mhq_streams_refpoints") %>%
  select(point_code, x, y)
mhq_streams_popunits <- read_vc(root =  "../data/v2024", file = "mhq_streams_popunits")

mhq_streams_assessments_recent <- mhq_streams_assessments %>%
  group_by(segment_id, point_code) %>%
  filter(assessment_date == max(assessment_date)) %>%
  ungroup() %>%
  left_join(mhq_streams_refpoints, by = "point_code") %>%
  left_join(select(mhq_streams_popunits, point_code, sac_streams, grts_ranking_draw), by = "point_code") %>%
  select(point_code, segment_id, grts_ranking_draw, sac_streams, x, y)

sample_unit_keep <- mhq_streams_assessments_recent %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  group_by(geometry) %>%
  mutate(dist_to_3260map = round(min(st_distance(map3260, geometry)), 1 )) %>%
  ungroup() %>%
  filter(!segment_id == "1_7_27869") #foutief punt
```



### Overlay measured sampling units with new sample frame

+ overlap 100 m segments with 10 meter buffer around existing sampling units
+ when 2 segments overlap with the same (buffer around a) sampling unit, select the segment with the highest rank (most upstream)

```{r}
watercourse_100mseg_overlap <- watercourse_100mseg_3260_lines %>%
  st_join(st_buffer(sample_unit_keep, dist = 10)) %>%
  filter(!is.na(segment_id)) 

# select the most upstream segment
max_ranking <- watercourse_100mseg_overlap %>%
  st_drop_geometry() %>%
  group_by(segment_id, point_code) %>%
  filter(rank == max(rank)) %>%
  ungroup()

watercourse_100mseg_overlap_unique <- watercourse_100mseg_overlap %>%
  filter(rank %in% max_ranking$rank)

sample_units_missing <- sample_unit_keep %>%
  filter(!segment_id %in% watercourse_100mseg_overlap$segment_id)
```

Possibly a segment starting from an existing sampling unit (point) will have a larger overlap with a more upstream watercourse segment (which it does not overlap with). 
Therefore we:

+ Include all upstream watercourse segment. 
+ For each watercourse segment point (the most downstream point) determine the distance to the closest existing sampling unit (point).

```{r}
watercourse_100mseg_3260_points_overlap <- watercourse_100mseg_3260_points %>%
  filter(rank %in% c(watercourse_100mseg_overlap_unique$rank, watercourse_100mseg_overlap_unique$rank + 1)) %>%
  st_join(select(sample_unit_keep, segment_id, point_code, sac_streams, grts_ranking_draw), join = st_nearest_feature) %>%
  group_by(geom) %>%
  mutate(dist = round(min(st_distance(geom, sample_unit_keep)), 1)) %>%
  ungroup()

```

In the map below we show:

+ a sampling unit of the original sample frame  (blue marker)
+ the overlapping 100 m segment in the new sample frame (blue line)
+ the next upstream 100 m segment (yellow line)
+ the 100 m segment points (orange circles)

In this case we will select the yellow line segment as it will have the higest overlap with the original sampling unit.

```{r}
sample_unit_orig <- sample_unit_keep %>%
  filter(segment_id == "1_4_1859") %>%
  st_transform(4326)

watercourse_select <- watercourse_100mseg_overlap %>%
  filter(segment_id == "1_4_1859") %>%
  st_transform(4326)

watercourse_point_select <- watercourse_100mseg_3260_points_overlap %>%
  filter(segment_id == "1_4_1859") %>%
  st_transform(4326)

watercourse_select2<- watercourse_100mseg_3260_lines %>%
  filter(rank == (watercourse_select$rank + 1)) %>%
  st_transform(4326)

sample_unit_orig %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers() %>%
  addCircleMarkers(data = watercourse_point_select, label =~rank, color = "orange") %>%
  addPolylines(data = watercourse_select, label = ~str_c(prop_3260, ";", rank)) %>%
  addPolylines(data = watercourse_select2, color = "yellow", label = ~str_c(prop_3260, ";", rank)) 
```

In the map below we will select the blue segment line.

```{r}
sample_unit_orig <- sample_unit_keep %>%
  filter(segment_id == "1_7_8845") %>%
  st_transform(4326)

watercourse_select <- watercourse_100mseg_overlap %>%
  filter(segment_id == "1_7_8845") %>%
  st_transform(4326)

watercourse_point_select <- watercourse_100mseg_3260_points_overlap %>%
  filter(segment_id == "1_7_8845") %>%
  st_transform(4326)

watercourse_select2 <- watercourse_100mseg_3260_lines %>%
  filter(rank == (watercourse_select$rank + 1)) %>%
  st_transform(4326)


sample_unit_orig %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers() %>%
  addCircleMarkers(data = watercourse_point_select, label =~rank, color = "orange") %>%
  addPolylines(data = watercourse_select, label = ~str_c(prop_3260, ";", rank)) %>%
  addPolylines(data = watercourse_select2, color = "yellow", label = ~str_c(prop_3260, ";", rank)) 
```

+ For each sampling unit select the closest watercourse segment point and corresponding watercourse segment line.

```{r}
watercourse_100mseg_existing <- watercourse_100mseg_3260_points_overlap %>%
  group_by(segment_id, point_code) %>%
  filter(dist == min(dist)) %>%
  ungroup() %>%
  mutate(sampleframe = "v1_7")

### add grts_ranking: for some sampling units location is replaced

# mhq_streams_invalid <- mhq_streams_refpoints_validity %>%
#   filter(!is_valid)
# 
# grts_ranking_orig <- mhq_streams_refpoints %>%
#   anti_join(mhq_streams_invalid, by = "point_code") %>%
#   select(segment_id, grts_ranking)
# 
# watercourse_100mseg_ranking_existing <- watercourse_100mseg_existing %>%
#   st_drop_geometry() %>%
#   select(rank, vhag_code, segment_id, sac_streams, grts_ranking_draw) %>%
#   left_join(grts_ranking_orig,
#             by = "segment_id") %>%
#   mutate(grts_ranking = ifelse(is.na(grts_ranking), grts_ranking_draw, grts_ranking),
#          sampleframe = "v1_7") 

watercourse_100mseg_ranking_existing <- watercourse_100mseg_existing
```

### Match other sampling units with new sample frame

+ First we simply match based on grts-ranking 

+ Then we select the watercourse segment lines that intersect with a buffer of 1 meter around the sampling units in the original sample frame

+ For each selected watercourse determine the nearest original sampling unit


```{r}

# these are the grts-rankings of the original sampling frame that have to be assigned to the sampling units of the new sampling frame
remaining_grts <- sample_frame_3260_v1_7 %>%
  st_drop_geometry() %>%
  filter(!grts_ranking %in% watercourse_100mseg_ranking_existing$grts_ranking_draw) %>%
  select(segment_id, grts_ranking) 

# these are the sampling units of the new sampling frame that occur in the original sampling frame
# grts_ranking = grts_ranking_draw
match_grts <- watercourse_100mseg_3260_ranking %>%
  semi_join(remaining_grts, by = c("grts_ranking")) %>%
  anti_join(watercourse_100mseg_ranking_existing, by = "rank") %>%
  group_by(grts_ranking) %>%
  mutate(existing_grts = TRUE,
         point_code = str_c(grts_ranking, "_", rank(rank))) %>%
  ungroup()

check <- match_grts %>%
  group_by(grts_ranking) %>%
  filter(n() > 1)

# the remaining grts-rankings that have not been assigned to the sampling units of the new sampling frame
remaining_grts <- sample_frame_3260_v1_7 %>%
  filter(!grts_ranking %in% watercourse_100mseg_ranking_existing$grts_ranking_draw) %>%
  filter(!grts_ranking %in% match_grts$grts_ranking)

# the sampling units of the new sampling frame (with rank as unique id ) that overlap with a buffer of 1 meter around the remaining grts-rankings in the original sampling frame (with segment_id as unique id)
  
watercourse_100mseg_overlap2 <- watercourse_100mseg_3260_lines %>%
  filter(!rank %in% watercourse_100mseg_ranking_existing$rank) %>%
  filter(!rank %in% match_grts$rank) %>%
  st_join(st_buffer(select(remaining_grts, segment_id, grts_ranking), dist = 1)) %>%
  filter(!is.na(segment_id)) %>%
  st_join(select(remaining_grts, segment_id_nearest = segment_id),
          join = st_nearest_feature) %>%
  mutate(check = segment_id == segment_id_nearest,
         point_code = str_c(grts_ranking, "_1"))

# when 2 su of the new sf overlap with the same su of the original sf, we select the su with the highest rank
# when 1 su of the new sf overlaps with more than 1 su of the original sf, we select the closest su
watercourse_100mseg_overlap2 <- watercourse_100mseg_overlap2 %>%
  st_drop_geometry() %>%
  group_by(segment_id) %>%
  filter(rank == max(rank)) %>%
  ungroup() %>%
  group_by(rank) %>%
  mutate(n_overlap = n_distinct(segment_id)) %>%
  ungroup() %>%
  filter(!((n_overlap > 1) & 
             (segment_id != segment_id_nearest))) %>%
  filter(!grts_ranking %in% watercourse_100mseg_ranking_existing$grts_ranking_draw) %>%
  filter(!grts_ranking %in% match_grts$grts_ranking) %>%
  mutate(existing_grts = FALSE)

check <- watercourse_100mseg_overlap2 %>%
  st_drop_geometry() %>%
  semi_join(watercourse_100mseg_ranking_existing, by = "point_code") 

watercourse_100mseg_overlap2 <- watercourse_100mseg_overlap2 %>%
  bind_rows(match_grts) %>%
  st_drop_geometry()
```

## Create sample frame

```{r}

# all matches between orignal su (segment_id) and new su (rank)
combine_grts_ranking_draw <- watercourse_100mseg_ranking_existing %>%
  bind_rows(watercourse_100mseg_overlap2 %>%
              rename(grts_ranking_draw = grts_ranking)) %>%
  st_drop_geometry() %>%
  select(rank, grts_ranking_draw, segment_id, point_code)

check <- combine_grts_ranking_draw %>%
  filter(n() > 1,
        .by = rank)

check <- combine_grts_ranking_draw %>%
  filter(n() > 1,
        .by = c(grts_ranking_draw, point_code))

# all matches between orignal su (segment_id) 
sampleframe_3260_ranking <- watercourse_100mseg_3260_ranking %>%
  mutate(sampleframe = "v2023") %>%
  left_join(combine_grts_ranking_draw, by = "rank") %>%
  mutate(grts_ranking_draw = ifelse(is.na(grts_ranking_draw), grts_ranking, grts_ranking_draw),
         new_point_code = is.na(point_code)) %>%
  group_by(grts_ranking_draw) %>%
  mutate(subrank = min_rank(pick(new_point_code, rank)), # create unique point codes
         n = n()) %>%
  ungroup() %>%
  mutate(point_code = ifelse(is.na(point_code), str_c(grts_ranking_draw, "_", subrank), point_code)) %>%
  select(rank, vhag_code, prop_3260, sac_streams, grts_ranking, grts_ranking_draw, segment_id, point_code, new_point_code) 

check <- sampleframe_3260_ranking %>%
  group_by(grts_ranking_draw , point_code) %>%
  filter(n() > 1)

check <- sampleframe_3260_ranking %>%
  group_by(point_code) %>%
  filter(n() > 1)

check_new_point_code <- sampleframe_3260_ranking %>%
  filter(new_point_code) %>%
  mutate(point_code_exists = point_code %in% mhq_streams_refpoints$point_code) %>%
  filter(point_code_exists)

# sampleframe_3260_ranking <- sampleframe_3260_ranking %>%
#   mutate(segment_id = ifelse(point_code == "1047854_1", "1_4_1088", segment_id)) # manual link

# sampleframe_3260_ranking <- sampleframe_3260_ranking %>%
#   left_join(select(mhq_streams_refpoints_valid, segment_id, point_code, db_ref), by = "segment_id") %>%
#   mutate(point_code = ifelse(is.na(point_code), str_c(grts_ranking_draw, "_1"), point_code),
#          point_code = ifelse(segment_id == "1_4_772" & !is.na(segment_id), str_c(grts_ranking_draw, "_2"), point_code ))
  
```


