
# Select extra sampling units for monitoring cycle 2

## Update mhq_streams_assessments

```{r, eval = FALSE}
mhq_streams_assessments_old <- read_vc(root =  "../../habitatstreams/output", file = "mhq_streams_assessments_v2022")

mhq_streams_assessments_new <- sample_cycle2_progress %>%
  left_join(select(mhq_streams_refpoints, segment_id, point_code, grts_ranking), by = "segment_id") %>%
  filter(!is.na(assessment_date)) %>%
  mutate(point_code = ifelse(is.na(point_code), str_c(grts_ranking_draw, "_1"), point_code),
         grts_ranking = ifelse(is.na(grts_ranking), grts_ranking_draw, grts_ranking), 
         point_code = ifelse(position_change == "ja", str_c(grts_ranking, "_2"), point_code)) %>%
  mutate(type_observed = ifelse(type_observed %in% c("gh", "3260"), type_observed, NA),
         inaccessible = ifelse(klasse == "uitval (ontoegankelijk)", "long term", NA),
         not_measurable = ifelse(klasse == "uitval (ongeschikt)", "long term",
                                 ifelse(klasse == "uitval (tijdelijk ongeschikt)", "short term", NA))) %>%
  select(point_code, segment_id, grts_ranking, grts_ranking_draw, assessment_date, type_observed, inaccessible, not_measurable, reason_no_assessment, x = x_assessment, y = y_assessment, x_orig = x, y_orig = y)

mq_streams_assessments_update <- mhq_streams_assessments_old %>%
  mutate(not_measurable = ifelse(!measurable, "long term", NA)) %>%
  filter(!is.na(assessment_date)) %>%
  bind_rows(mhq_streams_assessments_new) %>%
  select(point_code, segment_id, assessment_date, type_observed, inaccessible, not_measurable, x , y )

mq_streams_assessments_recent <- mq_streams_assessments_update %>%
  group_by(segment_id) %>%
  filter(assessment_date == max(assessment_date)) %>%
  ungroup()
```

```{r}
mhq_streams_assessments <- read_vc(root =  "../data/v2024", file = "mhq_streams_assessments")
mhq_streams_measurements <- read_vc(root =  "../data/v2024", file = "mhq_streams_measurements")
mhq_streams_popunits <- read_vc(root =  "../data/v2024", file = "mhq_streams_popunits")
mhq_streams_refpoints <- read_vc(root =  "../data/v2024", file = "mhq_streams_refpoints") %>%
  select(point_code, x, y)

mq_streams_assessments_recent <- mhq_streams_assessments %>%
  group_by(segment_id, point_code) %>%
  filter(assessment_date == max(assessment_date)) %>%
  ungroup()
```


## Select additional sampling units

We select 40 additional sampling unit per stratum based on grts_ranking_draw.

We include sampling units where habitat 3260 was not observed when:

+ the assessment was performed before 2019
+ the sampling unit is included in version 2023 of the 3260 habitat map 

For the sampling units that have been measured previously, we will select the original 100m segment point.
for the other sampling units we will select the 100 meter segment points and lines from the new sample frame.

```{r}
sampleframe_cycle2 <- sampleframe_3260_ranking %>%
  mutate(in_cycle2 = segment_id %in% sample_cycle2_progress$segment_id) %>%
  left_join(mq_streams_assessments_recent, by = c("point_code", "segment_id"))
```

```{r}
sample_add <- sampleframe_cycle2 %>%
  filter(prop_3260 > 10) %>%
  filter(point_code != "1047854_2") %>% #zeflde grtsranking als een bestaand meetpunt
  filter(!(in_cycle2 & year(assessment_date) >= 2022)) %>% #nog niet geselecteerd of nog niet bemonsterd in tweede cyclus
  filter(is.na(inaccessible) | inaccessible != "long term") %>% #toegankelijk
  filter(is.na(not_measurable) | not_measurable != "long term") %>% #geschikt
  filter(is.na(is_present) | is_present | ((!is_present) & year(assessment_date) < 2019)) %>%
  group_by(grts_ranking_draw) %>%
  mutate(double = n() > 1) %>%
  ungroup() %>%
  group_by(sac_streams) %>%
  slice_min(order_by = grts_ranking_draw, n = 40) %>%
  mutate(volgorde = rank(grts_ranking_draw)) %>%
  ungroup() %>%
  arrange(sac_streams, volgorde) 
```

+ For existing sampling units that have been measured we select the original 100 m segment

```{r}
sample_measured <- sample_add %>%
  rename(sampling_unit_code = segment_id) %>%
  filter(!is.na(assessment_date)) %>%
  filter(is_present) %>%
  mutate(type_meetpunt = "Bestaand meetpunt") %>%
  left_join(mhq_streams_refpoints, by = "point_code") %>%
  left_join(select(mhq_streams_popunits, point_code, db_ref), by = "point_code") %>%
  select(al_geselecteerd = in_cycle2, volgorde, sampling_unit_code, point_code, field_code = db_ref, grts_ranking, grts_ranking_draw, sac_streams, type_meetpunt, prop_3260, datum_opname_cyclus1 = assessment_date, x, y)

sample_measured_rank <- sample_add %>%
  filter(is_present) 

sample_measured_lines <- watercourse_100mseg_3260_lines %>%
  filter(rank %in% sample_measured_rank$rank)
  
```

+ For other sampling units we select the 100 m watercourse segments of the new sample frame.

```{r}

watercourse_100mseg_3260_coordinates <- watercourse_100mseg_3260_points %>%
  st_drop_geometry() %>%
  mutate(x = st_coordinates(watercourse_100mseg_3260_points)[,1],
         y = st_coordinates(watercourse_100mseg_3260_points)[,2]) %>%
  select(-vhag_code)

sample_not_measured <- sample_add %>%
  mutate(sampling_unit_code = str_c("2023_",rank)) %>%
  filter(is.na(assessment_date) | (!is_present)) %>%
  mutate(type_meetpunt = ifelse(is.na(is_present), "Nieuw meetpunt", "Meetpunt gh ouder dan 6 jaar")) %>%
  left_join(watercourse_100mseg_3260_coordinates, by = "rank") %>%
  left_join(select(mhq_streams_popunits, point_code, db_ref), by = "point_code") %>%
  select(al_geselecteerd = in_cycle2, volgorde, sampling_unit_code, point_code, field_code = db_ref, grts_ranking, grts_ranking_draw, sac_streams, type_meetpunt, prop_3260, datum_opname_cyclus1 = assessment_date, x, y) 

sample_not_measured_lines <- watercourse_100mseg_3260_lines %>%
  mutate(sampling_unit_code = str_c("2023_",rank)) %>%
  filter(sampling_unit_code %in% sample_not_measured$sampling_unit_code)
  
```

```{r}
sample_output <- sample_measured %>%
  bind_rows(sample_not_measured) %>%
  arrange(sac_streams, volgorde)
```

In the map below show

+ new sampling units (blue circle and blue line)
+ existing sampling units that have been measured previously (yellow circles) 
+ existing sampling units where 3260 habitat was not observed in an assessment before 2019 (red circles)

```{r}

sample_output %>%
  st_as_sf(coords = c("x", "y"), crs = 31370) %>%
  st_transform(4326) %>%
  mutate(show_color = ifelse(type_meetpunt == "Bestaand meetpunt", "yellow",
                             ifelse(type_meetpunt == "Nieuw meetpunt", "blue", "red"))) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(label = ~sampling_unit_code,
                   color = ~show_color) %>%
  addPolylines(data = st_transform(sample_not_measured_lines, 4326), label = ~sampling_unit_code) %>%
  addPolylines(data = st_transform(sample_measured_lines, 4326), label = ~rank, color = "blue")

```

## Output

```{r, eval = FALSE}
sample_output %>%
  write_csv2("../output/update_meetpunten3260_cyclus2_2025-07-15.csv")

sample_not_measured_lines %>%
  st_write("../output/meetpunten_100msegmenten_2025-07-15.shp", delete_layer = TRUE)
```
```{r}
sample_output_orig <- read_csv2("../output/update_meetpunten3260_cyclus2.csv")

check_verwijderd <- sample_output_orig %>%
  anti_join(sample_output, by = c("grts_ranking_draw", "sampling_unit_code")) %>%
  mutate(assessed = str_c(grts_ranking_draw, "_1") %in% mhq_streams_assessments$point_code,
         verandering = "verwijderd",
         reden = ifelse(assessed, "originele locatie van verplaats punt", 
                        ifelse(sampling_unit_code == "2023_98889", "nieuw punt dicht bij bestaand punt", "aangepaste selectie")))
  
check_toegevoegd <- sample_output %>%
  anti_join(sample_output_orig, by = "sampling_unit_code") %>%
  mutate(verandering = "toegevoegd") %>%
  select(-datum_opname_cyclus1)

verandering <- check_verwijderd %>%
  bind_rows(check_toegevoegd) %>%
  select(verandering, reden, sampling_unit_code, volgorde, field_code, sac_streams, type_meetpunt, prop_3260, x, y)

check <- sampleframe_cycle2 %>%
  semi_join(check_diff, by = "grts_ranking_draw") 

verandering %>%
  write_csv2("../output/wijziging_2025-07-15.csv")
  
```


## Update admin tables

### mhq streams refpoints

```{r}
admin_path <- fileman_up("n2khab-sample-admin")

refpoints_old <- read_vc(root =  "../data/v2024", file = "mhq_streams_refpoints")

refpoints_new <- sample_output %>%
  select(point_code, segment_id = sampling_unit_code, grts_ranking, type_meetpunt, x, y)

check <- refpoints_new %>%
  semi_join(refpoints_old, by = "point_code") %>%
  left_join(select(refpoints_old, point_code, segment_id_orig = segment_id, x_orig = x, y_orig = y), by = "point_code") %>%
  mutate(dist = sqrt((x - x_orig) ^ 2 + (y - y_orig) ^2) )

#OK

```

The table `mhq_streams_refpoints` defines the point location (the reference point) based on which the grts-ranking is assigned to a watercourse 100m segment.
Originally the reference points were always the most downstream points of each watercourse segment.
However, since we use the watercourse segments from the `watercourse_100mseg` this not always the case, since we matched the original reference points to the `watercourse_100mseg` segments.

`mhq_streams_refpoints` has following variables:

+ segment_id: the id of the watercourse segment
  + for watercourse segments that have been previously measured we use the segment_id from `samplingframe_habitatstreams_v1_7` 
  + for other watercourse segments we create a new id
+ rank: the `rank` (unique id) as defined in the `watercourse_100mseg` data source
  + when `rank` is missing, this is a watercourse segment in `samplingframe_habitatstreams_v1_7` that does not overlap with the `habitatstreams_2023`
+ segment_id_v1_7: the id of the watercourse segment in `samplingframe_habitatstreams_v1_7` based on which the reference point was originally defined
+ when segment_id_v1_7 is missing, the reference point corresponds with the most downstream point of the `watercourse_100mseg` segment
+ point_code: unique id for the reference point
+ x,y: coordinates (crs = 31370) 


```{r}
refpoints_new_coord <- sampleframe_cycle2 %>%
  filter(grts_ranking == grts_ranking_draw) %>%
  filter(is.na(assessment_date)) %>%
  left_join(watercourse_100mseg_3260_coordinates, by = "rank") %>%
  select(rank, segment_id, grts_ranking, grts_ranking_draw, sac_streams, prop_3260, point_code, in_cycle2, assessment_date, measured, x, y) %>%
  mutate(segment_id_new = str_c("2023_", rank))
  
refpoints_old_coord <- sampleframe_cycle2 %>%
  filter((grts_ranking != grts_ranking_draw) |
           !is.na(assessment_date)) %>%
  filter(str_sub(point_code, start = -1) == "1") %>%
  left_join(select(sample_frame_3260_v1_7_orig, segment_id, x, y), by = "segment_id") %>%
  select(rank, segment_id, grts_ranking, grts_ranking_draw, sac_streams, prop_3260, point_code, in_cycle2, assessment_date, measured, x, y) %>%
  mutate(grts_ranking = grts_ranking_draw,
         segment_id_v1_7 = segment_id,
         segment_id_new = ifelse(measured & !is.na(measured), segment_id, str_c("2023_", rank)))

refpoints_replaced_coord <- sampleframe_cycle2 %>%
  filter((grts_ranking != grts_ranking_draw) |
           !is.na(assessment_date)) %>%
  filter(str_sub(point_code, start = -1) == "2") %>%
  left_join(select(refpoints_old, point_code, x, y), by = "point_code") %>%
  select(rank, segment_id, grts_ranking, grts_ranking_draw, sac_streams, prop_3260, point_code, in_cycle2, assessment_date, measured, x, y) %>%
  mutate(segment_id_v1_7 = segment_id,
         segment_id_new = ifelse(measured & !is.na(measured), segment_id, str_c("2023_", rank)))

refpoints_replaced_coord <- refpoints_replaced_coord %>%
  st_as_sf(coords = c("x","y"), crs = 31370, remove = FALSE) 

refpoints_replaced_coord <- refpoints_replaced_coord %>%
  mutate(grts_ranking_check = (terra::extract(grts_master, refpoints_replaced_coord))$GRTSmaster_habitats) %>%
  mutate(grts_ranking = grts_ranking_check) %>%
  st_drop_geometry()
  
refpoints_old_keep <- refpoints_old %>%
  anti_join(refpoints_old_coord, by = "point_code") %>%
  anti_join(refpoints_replaced_coord, by = "point_code") %>%
  mutate(segment_id_new = segment_id,
         segment_id_v1_7 = segment_id)

# these are the assessed sampling units that do not overlap with the v2023 3260 habitatmap
check2 <- mhq_streams_assessments %>%
  anti_join(refpoints_old_coord, by = "point_code") %>%
  left_join(sample_units_missing, by = "point_code")
  
refpoints_update <- refpoints_new_coord %>%
  bind_rows(refpoints_old_coord) %>%
  bind_rows(refpoints_replaced_coord) %>%
  bind_rows(refpoints_old_keep) %>%
  select(segment_id = segment_id_new, rank, segment_id_v1_7, point_code, grts_ranking, x, y) %>%
  arrange(grts_ranking)

check <- refpoints_update %>%
  group_by(point_code) %>%
  filter(n() > 1)

check_ranking <- refpoints_update %>%
  st_as_sf(coords = c("x","y"), crs = 31370)

# grts-ranking of some replaced sampling units are not ok
check_ranking <- check_ranking %>%
  mutate(grts_ranking_check = (terra::extract(grts_master, check_ranking))$GRTSmaster_habitats) %>%
  filter(grts_ranking != grts_ranking_check)

```

```{r}
refpoints_update %>%
  write_vc(file = "mhq_streams_refpoints",
           root = "../output",
           digits = 20,
           sorting = c("grts_ranking", "point_code"))
```
### Population units

```{r}
mhq_streams_popunits_add <- sample_output %>%
  filter(!al_geselecteerd) %>%
  select(point_code, segment_id = sampling_unit_code, grts_ranking, grts_ranking_draw, sac_streams, db_ref = field_code) %>%
  mutate(grts_ranking = grts_ranking_draw,
         type = "3260",
         is_active = TRUE,
         legacy_site = FALSE,
         date = as.Date("2025-06-15"))
  
mhq_streams_popunits_update <- mhq_streams_popunits %>%
  bind_rows(mhq_streams_popunits_add) %>%
  group_by(point_code) %>%
  filter(date == max(date)) %>%
  ungroup()

check <- mhq_streams_popunits_update %>%
  group_by(point_code) %>%
  filter(n() > 1)
```

```{r}
mhq_streams_popunits_update %>%
  write_vc(file = "mhq_streams_popunits",
           root = "../output",
           digits = 20,
           sorting = c("grts_ranking_draw", "point_code"))
```

## Output sample frame

```{r}
sampleframe_3260 <- refpoints_new_coord %>%
  bind_rows(refpoints_old_coord) %>%
  bind_rows(refpoints_replaced_coord) %>%
  select(segment_id = segment_id_new, rank, segment_id_v1_7, prop_3260, sac_streams, point_code, grts_ranking, grts_ranking_draw) %>%
  arrange(sac_streams, grts_ranking_draw)

check <- sampleframe_3260 %>%
  filter(grts_ranking != grts_ranking_draw)
```

```{r}
sampleframe_3260 %>%
  write_vc(root = "../output", 
           file = "sampleframe_3260_grts_ranking",
           digits = 10,
           sorting = c("grts_ranking_draw", "rank"),
           strict = FALSE)
```