---
title: "Update overlay VBI and habitat map"
author: Toon Westra
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    code_folding: hide
    warning: FALSE
date: "2026-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
library(sf)
library(git2rdata)
library(RODBC)
library(leaflet)
library(n2khab)
library(kableExtra)
library(terra)
library(buffeRs)
library(crosstalk)
library(DT)
```

# Introduction

In 2024 we made an analysis of the most probable Natura 2000 forest habitat type for each sampling unit of the Flemish Forest Inventory (VBI).
This analysis was based on:

+ assessment of habitat type by VBI fieldworkers (only for part of the VBI sampling units)
+ habitatmap ([habitatmap_terr_2023_v1](https://zenodo.org/records/13886579))
+ manual control of habitat expert (when the observed habitat type was different than the type according to the habitatmap)

The file 'type_observed_vbi.csv' contains the results of the analysis and can be found in [this google drive folder](https://drive.google.com/drive/folders/1pqMzRD-wGZCjhgEfRMarRSQpazs9nVnR).

```{r}
vbi_habitatmap_2023 <- read_csv2("../output/type_observed_vbi.csv")

n_plots <- unique(vbi_habitatmap_2023$plot_id)
```

In 2026, the VBI monitoring protocol changes and vegetation cover and composition will only be assessed in sampling units that contain Natura 2000 habitat.
The file 'type_observed_vbi.csv' will be used to select the VBI sampling units with (high probability of) Natura 2000 habitat.

In 2025, a new habitat map was published: [habitatmap_2025](https://zenodo.org/records/18484399). 
We will update the file 'type_observed_vbi.csv' based on the new version of the habitat map.

# Sampling units from Flemish Forest inventory

We make use of following data sources (see [google drive](https://drive.google.com/drive/folders/1TjONf7WuoEYM6TzfHp9UW5KVw6ljMDgs)):

+ `vbi_strata_v2024-01-24.accdb`: observed habitat type (only available for the more recently measured plots)
+ `vbi_analysedatabank_v2023-08-22.accdb`: analysis variables per plot
+ `vbi_meetproces_v2022-11-15.accdb`: recorded coordinates

```{r}

db_strata <- "../data/vbi_strata_v2024-01-24.accdb"
db_meetproces <- "../data/vbi_meetproces_v2022-11-15.accdb"
db_analyse <- "../data/vbi_analysedatabank_v2023-08-22.accdb"

connect_db <-   odbcConnectAccess2007(db_strata)
strata_orig <- sqlFetch(connect_db, "tblplotstratadynamic")
odbcClose(connect_db)

connect_db <-   odbcConnectAccess2007(db_meetproces)
coordinaten_orig <- sqlFetch(connect_db, "tblcoordinaten")
odbcClose(connect_db)

connect_db <-   odbcConnectAccess2007(db_analyse)
check <- sqlTables(connect_db)
bestandskarakt_orig <- sqlFetch(connect_db, "tbl2BestandskaraktKwal")
cover_species_vbi <- sqlFetch(connect_db, "tbl11VegetatieSamenstelling")
measurements_vbi <- sqlFetch(connect_db, "tbl4Vegetatie")
odbcClose(connect_db)

strata <- strata_orig %>%
  select(plot_id = IDPlots, segment_id = IDSegments, periode = Periode, landuse = LandUse, stand_type = StandType, type_observed = Habitat) %>%
  filter(!is.na(stand_type)) %>%
  select(-stand_type)

coordinaten <- coordinaten_orig %>%
  select(plot_id = idplots, x, y, type_coord = type, periode, reeks, plotweight)

bestandskarakt <- bestandskarakt_orig %>%
  select(plot_id = IDPlots, point_id = IDGroup, periode = Periode, year = DateDendro, segment_id = IDSegments, stand_type = v2_StandType)

measurements_vbi <- measurements_vbi %>%
  select(plot_id = IDPlots, segment_id = IDSegments, periode = Periode, veg_record = VegRecord, dendro_record = DendroRecord, dendro_record_zero = DendroRecordZero)

cover_species_vbi <- cover_species_vbi %>%
  select(plot_id = IDPlots, periode = Periode, layer = Layer, name_nl = NameNl, name_sc = NameSc, cover_categ =  Cover_categ, date_veg = DateVeg)

coord_recent <- coordinaten %>%
  semi_join(cover_species_vbi, by = c("plot_id")) %>%
  group_by(plot_id) %>%
  slice_max(periode) %>%
  ungroup()

#write_csv2(coord_recent, "../output/coord_vbi.csv")
```

```{r}
#we willen reeks toevoegen ivm planning
vbi_reeks <- coordinaten_orig %>%
  group_by(idplots) %>%
  filter(periode == max(periode)) %>%
  ungroup() %>%
  distinct(idplots, reeks, x, y) %>%
  rename(plot_id = idplots)

plot_reeks <- vbi_reeks %>%
  distinct(plot_id, reeks)

#meest recente opname per meetpunt
bestandskarakt_actual <- bestandskarakt %>%
  group_by(plot_id) %>%
  slice_max(periode) %>%
  ungroup() %>%
  filter(segment_id == 1) %>%
  filter(periode > 1) %>%
  filter(!is.na(stand_type))

vbi_meetpunten_sf <- bestandskarakt_actual %>%
  left_join(strata, by = c("plot_id", "periode", "segment_id")) %>%
  left_join(coordinaten, by = c("plot_id", "periode")) %>%
  st_as_sf(coords = c("x", "y"), crs = 31370)

overzicht <- vbi_meetpunten_sf %>%
  st_drop_geometry() %>%
  mutate(habtype_assessed = !is.na(type_observed)) %>%
  group_by(periode, reeks, habtype_assessed) %>%
  summarise(year_min = min(year, na.rm = TRUE),
            year_max = max(year, na.rm = TRUE),
            n = n_distinct(point_id)) %>%
  ungroup()
```

For each sampling unit we select the most recent measurement (available in the dataset) and the corresponding coordinates.

The table below gives an overview of the number of VBI plots for which the observed habitat type was assessed.

```{r}
overzicht %>%
  datatable(rownames = FALSE,
            filter = "top")
```

# New habitat map

`habitatmap_terr_2025_v1` is not yet published.
We use a temporary file.
To do: publish data source and add link.

```{r}
habitatmap_2025 <- read_habitatmap_stdized()
```

# Overlay VBI sampling units and habitatmap2025

+ Overlay between measured VBI sampling units and the habitatmap2025.
+ Selection of sampling units that are located in habitatmap2025 polygon with habitat type 2180 or 91xx
+ Selection of sampling units for which the observed habitat type has not been assessed yet (the other sampling units have already been checked by the habitat expert)

```{r}
vbi_meetpunten_sf <- vbi_meetpunten_sf %>%
  st_join(select(habitatmap_2025$habitatmap_polygons, polygon_id))

```

```{r}
vbi_meetpunten_hab <- vbi_meetpunten_sf %>%
  st_drop_geometry() %>%
  left_join(habitatmap_2025$habitatmap_types, by = "polygon_id") %>%
  mutate(main_type = str_sub(type, 1, 4)) %>%
  left_join(measurements_vbi, by = c("plot_id", "segment_id", "periode")) %>%
  mutate(measured = veg_record & (is.na(dendro_record) | (dendro_record | dendro_record_zero)))

vbi_meetpunten_hab_update <- vbi_meetpunten_hab %>%
  filter(is.na(type_observed) | type_observed %in% c("9 : bossen", "91E0", "91E0*")) %>%
  filter(!is.na(polygon_id)) %>%
  filter(str_detect(type, "91|2180")) %>%
  filter(certain | (type == "91E0")) %>%
  mutate(new_plot = !plot_id %in% vbi_habitatmap_2023$plot_id)

check_new <- vbi_meetpunten_hab_update %>%
  filter(new_plot)

check_removed <- vbi_habitatmap_2023 %>%
  filter(source == "habitatmap2023") %>%
  filter(!plot_id %in% vbi_meetpunten_hab_update$plot_id) %>%
  left_join(select(vbi_meetpunten_hab, plot_id, reeks, type_observed, polygon_id, type_2025 = type, stand_type, landuse, certain, code_orig), by = c("plot_id"))
```

Following sampling units are new compared to the previous selection (i.e. `vbi_type_observed.csv`) based on habitatmap2023.

```{r}
check_new %>%
  select(plot_id, point_id, reeks, year_measured = year, stand_type, type, phab) %>%
  arrange(reeks, plot_id) %>%
  datatable(rownames = FALSE,
            filter = "top")
```
Following sampling units are removed compared to the previous selection (i.e. `vbi_type_observed.csv`) based on habitatmap2023.

```{r}
check_removed %>%
  select(plot_id, reeks, source, stand_type, type, phab, type_2025, polygon_id) %>%
  arrange(reeks, plot_id) %>%
  datatable(rownames = FALSE,
            filter = "top")
```

# Update table `type_observed_vbi`

+ From the previous version we remove all records for which `source = habitatmap2023`.
+ We add the new records based on habitatmap2025.

We write the result to the file 'type_observed_vbi_2025.csv'.

```{r}
type_observed_vbi_update <- vbi_habitatmap_2023 %>%
  left_join(plot_reeks, by = "plot_id") %>%
  filter(source != "habitatmap2023") %>%
  bind_rows(vbi_meetpunten_hab_update %>%
              mutate(source = "habitatmap2025") %>%
              select(plot_id, reeks, periode, source, type, phab)) %>%
  arrange(reeks, plot_id) %>%
  select(plot_id, reeks, periode, type, phab, source, note_manual_check)

n_plots <- type_observed_vbi_update$plot_id %>%
  unique()

# type_observed_vbi_update %>%
#   write_csv2("../output/type_observed_vbi_2025.csv")
```

# Future updates

The table will be updated in the future when:

+ a dataset with new VBI habitat type assessments becomes available
+ a new version of the habitatmap becomes available

